### 2019-05-13 report/report.Rmd
Initial report template with authors in header

---
title: '<header style="text-align: center">
			<h1 style="font-weight: bold; font-size: 44px">2016 Brussels Bombings</h1>
			<h2>Air Traffic Analysis</h2>
		</header>'
author: '<div style="text-align: center">
			<span style="display: inline-block; width: 30%; margin: 0 5px">[Ramiro Balado Ordax](https://github.com/Qjammer)</span>
			<span style="display: inline-block; width: 30%; margin: 0 5px">[Alejandro Hernández Rubio](https://github.com/Shyryuu)</span>
			<span style="display: inline-block; width: 30%; margin: 0 5px">[Pol Mesalles Ripoll](https://github.com/polmes)</span>
		</div>'
# date: "DD/MM/YYYY"
output:
    rmarkdown::html_document:
        theme: cosmo
        highlight: tango
        number_sections: true
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-05-13 report/report.Rmd
Skeleton report structure

@@
 
-## R Markdown
+# Air Traffic
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+# Airports
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+# Airlines

### 2019-05-15 main.R
Finished processing algorithm
Processing of *.so6 data files should be complete now:
- Adds airline codes
- Removes technical and other unnecessary points (when possible)
- Returns list with tables:
	- FLIGHTS: id orig dest callsign airline aircraft cumdist takeoff landing
	- ROUTES: id dt lat lon FL
	- AIRPLANES: callsign airline aircraft mindate maxdate

@@ -2,16 +2,28 @@
 requirements <- c('data.table', 'fasttime', 'ggplot2')
 
 # Check requirements
-missing <- requirements[!(requirements %in% installed.packages()[,'Package'])]
+missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]
 if (length(missing)) {
 	message('Installing missing packages...')
 	install.packages(missing)
 }
 
 # Load requirements
-slibrary <- function(pkg) { suppressPackageStartupMessages(library(pkg, character.only=TRUE, quietly=TRUE)) }
+slibrary <- function(pkg) { suppressPackageStartupMessages(library(pkg, character.only = TRUE, quietly = TRUE)) }
 lapply(requirements, slibrary)
 
 # Set working directory to source directory
 sourceDir <- getSrcDirectory(function() {}) # R is dumb
 setwd(sourceDir)
+
+# if RDS exists:
+	# load
+# else:
+	# process data/ files
+	# ls <- list.files(pattern = '\\.so6$')
+
+source('processor.R')
+file <- 'data/20160222_20160226_0000_2359_____m3.so6'
+system.time(
+	test <- processor(file, getairplanes = TRUE)
+)

### 2019-05-15 processor.R

@@ -1,57 +1,98 @@
-library(data.table)
-library(fasttime)
-tables<-function(file){
-	colnames<-c("SegID","ICAO_orig","ICAO_dest","AC_type",
-				"time1","time2","FL1","FL2",
-				"Callsign","Date_st","Date_end",
-				"Lat_st","Lon_st","Lat_end","Lon_end","FlID","seq","Dist")
-	#colClasses<-list(factor=c("Time1","Time2","Date1","Date2"))
-	colClasses<-list(character=c(5,6,11,12))
-	message("Reading File")
-	ddr2<-fread(file,drop=c(9,20),stringsAsFactors=TRUE,col.names=colnames,colClasses=colClasses)
-	ddr2<-ddr2[!ICAO_orig %in% c("ZZZZ","AFIL") & !ICAO_dest %in% c("ZZZZ","AFIL")]
+processor <- function(file, getairplanes = FALSE) {
+	# Extract filename from path
+	filename <- basename(file)
+	message(paste('Processing file: ', filename))
+	
+	# m1 = plan / m3 = real
+	isreal <- if (substr(file, nchar(file) - 4, nchar(file) - 4) == '3') TRUE else FALSE
+	
+	colnames <- c('pts', 'orig', 'dest', 'aircraft', 'time1', 'time2', 'FL1', 'FL2', 'callsign',
+				  'date1', 'date2', 'lat1', 'lon1', 'lat2', 'lon2', 'id', 'dist')
+	colclasses <- list(character = c(5,6,11,12)) # time1, time2, date1, date2
+	
+	message('Reading file...')
+	ddr <- fread(file, drop = c(9, 18, 20), stringsAsFactors = TRUE, col.names = colnames, colClasses = colclasses)
+	format(object.size(ddr), units = 'MiB')
 
-	message("Splitting IDs")
-	ddr2[, c("PtID_st", "PtID_end") := tstrsplit(SegID, "_", fixed = TRUE)]
-	message("Converting to dates")
-	ddr2[, Date_st := fastPOSIXct(paste(substr(Date_st,1,2),"-",substr(Date_st,3,4),"-",substr(Date_st,5,6),"T", substr(time1,1,2),":",substr(time1,3,4),":",substr(time1,5,6)))]
[…]

### 2019-05-17 processor.R
Fix lat/lon values dividing by 60

@@ -2,14 +2,14 @@ processor <- function(file, getairplanes = FALSE) {
 	# Extract filename from path
 	filename <- basename(file)
 	message(paste('Processing file: ', filename))
-	
+
 	# m1 = plan / m3 = real
 	isreal <- if (substr(file, nchar(file) - 4, nchar(file) - 4) == '3') TRUE else FALSE
-	
+
 	colnames <- c('pts', 'orig', 'dest', 'aircraft', 'time1', 'time2', 'FL1', 'FL2', 'callsign',
 				  'date1', 'date2', 'lat1', 'lon1', 'lat2', 'lon2', 'id', 'dist')
 	colclasses <- list(character = c(5,6,11,12)) # time1, time2, date1, date2
-	
+
 	message('Reading file...')
 	ddr <- fread(file, drop = c(9, 18, 20), stringsAsFactors = TRUE, col.names = colnames, colClasses = colclasses)
 	format(object.size(ddr), units = 'MiB')
@@ -17,7 +17,7 @@ processor <- function(file, getairplanes = FALSE) {
 	message('Removing flights with special origins or destinations...')
 	ddr <- ddr[!orig %in% c('ZZZZ', 'AFIL') & !dest %in% c('ZZZZ', 'AFIL')]
 	format(object.size(ddr), units = 'MiB')
-	
+
 	message('Removing non-commercial traffic...')
 	airlines <- readRDS('data/airlines.RDS')
 	availablemonths <- unique(airlines$month)
@@ -29,11 +29,11 @@ processor <- function(file, getairplanes = FALSE) {
 	ddr <- ddr[airline %in% airlines$ICAO]
 	rm(airlines)
 	format(object.size(ddr), units = 'MiB')
-	
+
 	message('Computing cumulative distance...')
 	ddr[, cumdist := cumsum(dist), by = id]
 	format(object.size(ddr), units = 'MiB')
-	
+
 	# message('Removing technical-technical pairs of points...')
 	# ddr <- ddr[!grepl('^[$%#!].+_[$%#!]', ddr$pts),]
 	# format(object.size(ddr), units = 'MiB')
@@ -41,7 +41,7 @@ processor <- function(file, getairplanes = FALSE) {
 	# message('Splitting IDs...')
 	# ddr[, c('pt1', 'pt2') := tstrsplit(pts, '_', fixed = TRUE)]
 	# format(object.size(ddr), units = 'MiB')
-	
+
[…]

### 2019-05-17 util/process.R
Rename processor to ddr.process and move to 'util/'

@@ -1,4 +1,4 @@
-processor <- function(file, getairplanes = FALSE) {
+ddr.process <- function(file, getairplanes = FALSE) {
 	# Extract filename from path
 	filename <- basename(file)
 	message(paste('Processing file: ', filename))

### 2019-05-18 util/map.R
Add map plotting function
Includes autocenter parameter to focus on center of data or center of Europe

@@ -1,0 +1,28 @@
+ddr.map <- function(routes, autocenter = TRUE) {
+	map <- ggplot(map_data('world'), aes(x = long, y = lat, group = group)) +
+		geom_polygon(fill = 'white', color = 'black', size = 0.2) +
+		geom_line(data = routes,
+				  aes(x = lon, y = lat, group = id, color = rainbow(routes[, .N])),
+				  lwd = 0.25) +
+		theme(legend.position = 'none',
+			axis.line = element_blank(), axis.ticks = element_blank(),
+			axis.text.x = element_blank(), axis.text.y = element_blank(),
+			axis.title.x = element_blank(), axis.title.y = element_blank(),
+			panel.border = element_rect(fill = NA, color = 'black')
+		) +
+		if (autocenter) {
+			coord_map('ortho',
+					  orientation = c(median(routes[, lat], na.rm = TRUE), median(routes[, lon], na.rm = TRUE), 0),
+					  xlim = c(min(routes[, lon], na.rm = TRUE) - 2, max(routes[, lon], na.rm = TRUE) + 2),
+					  ylim = c(min(routes[, lat], na.rm = TRUE) - 2, max(routes[, lat], na.rm = TRUE) + 2)
+			)
+		} else {
+			coord_map('ortho',
+					  orientation = c(50.8, 4.4, 0), # Brussels
+					  xlim = c(-20, 50), ylim = c(32, 70) # Europe
+			)
+		}
+
+	print(map)
+	return(map)
+}

### 2019-05-18 util/process.R
Remove 'isreal' parameter from single tables

@@ -1,10 +1,10 @@
 ddr.process <- function(file, getairplanes = FALSE) {
 	# Extract filename from path
 	filename <- basename(file)
-	message(paste('Processing file: ', filename))
+	message(paste('Processing file:', filename))
 
 	# m1 = plan / m3 = real
-	isreal <- if (substr(file, nchar(file) - 4, nchar(file) - 4) == '3') TRUE else FALSE
+	# isreal <- if (substr(file, nchar(file) - 4, nchar(file) - 4) == '3') TRUE else FALSE
 
 	colnames <- c('pts', 'orig', 'dest', 'aircraft', 'time1', 'time2', 'FL1', 'FL2', 'callsign',
 				  'date1', 'date2', 'lat1', 'lon1', 'lat2', 'lon2', 'id', 'dist')
@@ -19,7 +19,7 @@ ddr.process <- function(file, getairplanes = FALSE) {
 	format(object.size(ddr), units = 'MiB')
 
 	message('Removing non-commercial traffic...')
-	airlines <- readRDS('data/airlines.RDS')
+	airlines <- readRDS(ddr.path('airlines'))
 	availablemonths <- unique(airlines$month)
 	filedate <- as.Date(substr(filename, 1, 8), '%Y%m%d')
 	filemonth <- availablemonths[which.min(abs(availablemonths - filedate))]
@@ -55,6 +55,7 @@ ddr.process <- function(file, getairplanes = FALSE) {
 	flights <- flights[, .SD[.N], by = id]
 	flights[, takeoff := ddr[, .SD[1], by = id, .SDcols = c('dt1')][, .(dt1)]]
 	flights[, landing := ddr[, .SD[.N], by = id, .SDcols = c('dt2')][, .(dt2)]]
+	# flights[, isreal := isreal]
 	flights <- unique(flights)
 	format(object.size(flights), units = 'MiB')
 
@@ -78,7 +79,7 @@ ddr.process <- function(file, getairplanes = FALSE) {
 	if (!getairplanes) {
 		rm(ddr)
 
-		tb <- list(isreal = isreal, flights = flights, routes = routes)
+		tb <- list(flights = flights, routes = routes)
 	} else {
 		# AIRPLANES: callsign airline aircraft mindate maxdate
 		message('Processing airplane data...')
@@ -89,8 +90,9 @@ ddr.process <- function(file, getairplanes = FALSE) {
 		airplanes <- unique(airplanes)
[…]

### 2019-05-18 util/load.R
Add function to load *.RDS or process *.so6 files
It will be used in main, called within 'lapply'; the function saves processed files to *.RDS for future use

@@ -1,0 +1,25 @@
+ddr.load <- function(name) {
+	file <- ddr.path(name)
+	if (file.exists(file)) {
+		# Load *.RDS file
+		message(paste('Loading file:', basename(file)))
+		datedata <- readRDS(file) # includes both real and plan data
+	} else {
+		# Get full *.so6 filenames
+		both <- list.files(path = 'data', pattern = paste0('^', name, '.*\\.so6$'), full.names = TRUE)
+		if (length(both) != 2) stop('Did not find expected files')
+
+		# Process both *.so6 files
+		realdata <- ddr.process(both[grepl('m3', both)]) # m3 = real
+		plandata <- ddr.process(both[grepl('m1', both)]) # m1 = plan
+
+		# Join real and plan data
+		datedata <- list(real = realdata, plan = plandata)
+
+		# Save for next time
+		filename <- ddr.path(name)
+		message(paste('Saving to file:', basename(filename)))
+		saveRDS(datedata, filename)
+	}
+	return(datedata)
+}

### 2019-05-18 main.R
Finished data merging
The code now works with multiple data files, which are combined together
(with 'rbindlist') to create a single list of lists with the following
structure:
- real
	- flights
	- routes
- plan
	- flights
	- routes
The resulting data file is saved as *.RDS for future use

@@ -1,5 +1,11 @@
+# Input date range
+# day1 <- '2016/02/22'
+# day2 <- '2016/03/02'
+day1 <- '2019/02/08'
+day2 <- '2019/02/13'
+
 # Define required packages
-requirements <- c('data.table', 'fasttime', 'ggplot2')
+requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj')
 
 # Check requirements
 missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]
@@ -9,21 +15,53 @@ if (length(missing)) {
 }
 
 # Load requirements
-slibrary <- function(pkg) { suppressPackageStartupMessages(library(pkg, character.only = TRUE, quietly = TRUE)) }
+slibrary <- function(pkg) suppressPackageStartupMessages(library(pkg, character.only = TRUE, quietly = TRUE))
 lapply(requirements, slibrary)
 
 # Set working directory to source directory
 sourceDir <- getSrcDirectory(function() {}) # R is dumb
 setwd(sourceDir)
 
-# if RDS exists:
-	# load
-# else:
-	# process data/ files
-	# ls <- list.files(pattern = '\\.so6$')
+# Load functions in 'util/'
+lapply(list.files(path = 'util', full.names = TRUE), source)
 
-source('processor.R')
-file <- 'data/20160222_20160226_0000_2359_____m3.so6'
-system.time(
-	test <- processor(file, getairplanes = TRUE)
-)
+# Get file path util function
+ddr.path <- function(name, ext = '.RDS') return(file.path('data', paste0(name, ext)))
+
+# Load or process 'data/' files
+day1 <- as.Date(day1, '%Y/%m/%d')
+day2 <- as.Date(day2, '%Y/%m/%d')
+
+name <- paste0(format(day1, '%Y%m%d'), '_', format(day2, '%Y%m%d'))
+file <- ddr.path(name)
+if (file.exists(file)) {
+	message(paste('Loading full file:', basename(file)))
+	data <- readRDS(file)
+} else {
+	# Divide in intervals of 5 days
+	days <- seq(day1, day2, by = 'days')
+	int1 <- days[seq(1, length(days), 5)]
+	int2 <- if (length(days) >= 5) days[seq(5, length(days), 5)] else day2
+	names <- paste0(format(int1, '%Y%m%d'), '_', format(unique(c(int2, day2)), '%Y%m%d'))
+
[…]

### 2019-05-18 main.R
Test map plotting

@@ -1,5 +1,5 @@
 # Define required packages
-requirements <- c('data.table', 'fasttime', 'ggplot2')
+requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj')
 
 # Check requirements
 missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]
@@ -16,14 +16,18 @@ lapply(requirements, slibrary)
 sourceDir <- getSrcDirectory(function() {}) # R is dumb
 setwd(sourceDir)
 
+# Load functions in 'util/'
+lapply(list.files(path = 'util', full.names = TRUE), source)
+
 # if RDS exists:
 	# load
 # else:
 	# process data/ files
 	# ls <- list.files(pattern = '\\.so6$')
 
-source('processor.R')
+# Test data processing
 file <- 'data/20160222_20160226_0000_2359_____m3.so6'
-system.time(
-	test <- processor(file, getairplanes = TRUE)
-)
+test <- ddr.process(file, getairplanes = TRUE)
+
+# Test map plotting
+ddr.map(test$routes[1:1000], autocenter = FALSE)

### 2019-05-18 main.R
Changed input dates back to Brussels case study

@@ -1,8 +1,6 @@
 # Input date range
-# day1 <- '2016/02/22'
-# day2 <- '2016/03/02'
-day1 <- '2019/02/08'
-day2 <- '2019/02/13'
+day1 <- '2016/02/22'
+day2 <- '2016/03/02'
 
 # Define required packages
 requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj')
@@ -59,9 +57,5 @@ if (file.exists(file)) {
 	saveRDS(data, file)
 }
 
-# Test data processing
-# file <- 'data/20160222_20160226_0000_2359_____m3.so6'
-# test <- ddr.process(file, getairplanes = TRUE)
-#
-# # Test map plotting
-# ddr.map(test$routes[1:1000], autocenter = FALSE)
+# Test map plotting
+ddr.map(data$real$routes[1:1000], autocenter = FALSE)

### 2019-05-18 util/process.R
Actually save 'pt' column
We need to keep the point id column to find diverted flights, maybe more

@@ -57,22 +57,25 @@ ddr.process <- function(file, getairplanes = FALSE) {
 	flights[, landing := ddr[, .SD[.N], by = id, .SDcols = c('dt2')][, .(dt2)]]
 	# flights[, isreal := isreal]
 	flights <- unique(flights)
+	setorder(flights, id)
 	format(object.size(flights), units = 'MiB')
 
 	# ROUTES: id dt lat lon FL
 	message('Processing route (trajectory) data...')
-	start <- ddr[, .SD[1], by = id, .SDcols = c('dt1', 'lat1', 'lon1', 'FL1')]
+	start <- ddr[, .SD[1], by = id, .SDcols = c('pts', 'dt1', 'lat1', 'lon1', 'FL1')]
+	start[, pts := sub('^(.+)_.*$', '\\1', pts)]
 	ddr[, c('lat1', 'lon1', 'FL1') := NULL]
-	rest <- ddr[!grepl('_[$%#!]', ddr$pts), c('id', 'dt2', 'lat2', 'lon2', 'FL2')]
+	rest <- ddr[!grepl('_[$%#!]', ddr$pts), c('id', 'pts', 'dt2', 'lat2', 'lon2', 'FL2')]
+	rest[, pts := sub('^.*_(.+)$', '\\1', pts)]
 	ddr[, c('pts', 'lat2', 'lon2', 'FL2') := NULL]
-	routecols <- c('id', 'dt', 'lat', 'lon', 'FL')
+	routecols <- c('id', 'pt', 'dt', 'lat', 'lon', 'FL')
 	setnames(start, colnames(start), routecols)
 	setnames(rest, colnames(rest), routecols)
 	routes <- rbindlist(list(start, rest))
 	routes[, `:=` (lat = lat/60, lon = lon/60)]
 	rm(start, rest)
-	setorder(routes, id, dt)
 	routes <- unique(routes)
+	setorder(routes, id, dt)
 	format(object.size(routes), units = 'MiB')
 
 	# Export list of tables

### 2019-05-18 util/process.R
Remove unused format statements that printed variable size

@@ -12,11 +12,9 @@ ddr.process <- function(file, getairplanes = FALSE) {
 
 	message('Reading file...')
 	ddr <- fread(file, drop = c(9, 18, 20), stringsAsFactors = TRUE, col.names = colnames, colClasses = colclasses)
-	format(object.size(ddr), units = 'MiB')
 
 	message('Removing flights with special origins or destinations...')
 	ddr <- ddr[!orig %in% c('ZZZZ', 'AFIL') & !dest %in% c('ZZZZ', 'AFIL')]
-	format(object.size(ddr), units = 'MiB')
 
 	message('Removing non-commercial traffic...')
 	airlines <- readRDS(ddr.path('airlines'))
@@ -28,25 +26,20 @@ ddr.process <- function(file, getairplanes = FALSE) {
 	ddr[grepl('^[[:upper:]]{3}[[:digit:]]', ddr$callsign), airline := substr(callsign, 1, 3)]
 	ddr <- ddr[airline %in% airlines$ICAO]
 	rm(airlines)
-	format(object.size(ddr), units = 'MiB')
 
 	message('Computing cumulative distance...')
 	ddr[, cumdist := cumsum(dist), by = id]
-	format(object.size(ddr), units = 'MiB')
 
 	# message('Removing technical-technical pairs of points...')
 	# ddr <- ddr[!grepl('^[$%#!].+_[$%#!]', ddr$pts),]
-	# format(object.size(ddr), units = 'MiB')
 
 	# message('Splitting IDs...')
 	# ddr[, c('pt1', 'pt2') := tstrsplit(pts, '_', fixed = TRUE)]
-	# format(object.size(ddr), units = 'MiB')
 
 	message('Converting dates...')
 	ddr[, dt1 := fastPOSIXct(paste(substr(date1,1,2),'-',substr(date1,3,4),'-',substr(date1,5,6),'T', substr(time1,1,2),':',substr(time1,3,4),':',substr(time1,5,6)))]
 	ddr[, dt2 := fastPOSIXct(paste(substr(date2,1,2),'-',substr(date2,3,4),'-',substr(date2,5,6),'T', substr(time2,1,2),':',substr(time2,3,4),':',substr(time2,5,6)))]
 	ddr[, c('date1', 'date2', 'time1', 'time2') := NULL]
-	format(object.size(ddr), units = 'MiB')
 
 	# FLIGHTS: id orig dest callsign airline aircraft cumdist takeoff landing
 	message('Processing flight data...')
[…]

### 2019-05-19 util/map.R
Fix code formatting issues

@@ -8,19 +8,16 @@ ddr.map <- function(routes, autocenter = TRUE) {
 			axis.line = element_blank(), axis.ticks = element_blank(),
 			axis.text.x = element_blank(), axis.text.y = element_blank(),
 			axis.title.x = element_blank(), axis.title.y = element_blank(),
-			panel.border = element_rect(fill = NA, color = 'black')
-		) +
+			panel.border = element_rect(fill = NA, color = 'black')) +
 		if (autocenter) {
 			coord_map('ortho',
 					  orientation = c(median(routes[, lat], na.rm = TRUE), median(routes[, lon], na.rm = TRUE), 0),
 					  xlim = c(min(routes[, lon], na.rm = TRUE) - 2, max(routes[, lon], na.rm = TRUE) + 2),
-					  ylim = c(min(routes[, lat], na.rm = TRUE) - 2, max(routes[, lat], na.rm = TRUE) + 2)
-			)
+					  ylim = c(min(routes[, lat], na.rm = TRUE) - 2, max(routes[, lat], na.rm = TRUE) + 2))
 		} else {
 			coord_map('ortho',
 					  orientation = c(50.8, 4.4, 0), # Brussels
-					  xlim = c(-20, 50), ylim = c(32, 70) # Europe
-			)
+					  xlim = c(-20, 50), ylim = c(32, 70)) # Europe
 		}
 
 	print(map)

### 2019-05-19 util/times.R
Add interactive time series plot function

@@ -1,0 +1,38 @@
+ddr.times <- function(flights, airport, event, onlydest = FALSE) {
+	cols <- c('date', 'N')
+	flightsWithDest <- flights[dest == airport, .N,  by = as.Date(landing)]
+	setnames(flightsWithDest, cols)
+
+	if (!onlydest) {
+		flightsWithOrig <- flights[orig == airport, .N,  by = as.Date(takeoff)]
+		setnames(flightsWithOrig, cols)
+		allFlights <- merge(flightsWithOrig, flightsWithDest, by = c('date'))
+	} else {
+		allFlights <- flightsWithDest
+	}
+
+	if (!onlydest) {
+		dygraph(allFlights,
+				main = paste('Air Traffic Evolution at', airport),
+				ylab = 'Number of Flights',
+				width = '100%') %>%
+			dySeries('N.x', label = paste(airport, 'as Origin')) %>%
+			dySeries('N.y', label = paste(airport, 'as Destination')) %>%
+			dyRangeSelector(dateWindow = event) %>%
+			dyOptions(colors = c('royalblue', 'coral')) %>%
+			dyLegend(width = '500') %>%
+			dyEvent(event[1], 'Disruption begins', labelLoc = 'bottom') %>%
+			dyEvent(event[2], 'Disruption ends', labelLoc = 'bottom')
+	} else {
+		dygraph(allFlights,
+				main = paste('Air Traffic Evolution at', airport),
+				ylab = paste('Number of Flights')) %>%
+			dySeries('N', label = paste(airport, 'as Destination')) %>%
+			dyRangeSelector(dateWindow = event) %>%
+			dyOptions(colors = c('royalblue', 'coral')) %>%
+			dyLegend(show = 'always', width = '500') %>%
+			dyEvent(event[1], 'Disruption begins', labelLoc = 'bottom') %>%
+			dyEvent(event[2], 'Disruption ends', labelLoc = 'bottom')
+	}
+
+}

### 2019-05-19 util/render.R
Render function that calls rmarkdown::render and opens result in viewer

@@ -1,0 +1,6 @@
+ddr.render <- function(file, dir = 'report') {
+	rep <- rmarkdown::render(file.path(dir, file), quiet = TRUE)
+	tmp <- tempfile(file, fileext = '.html')
+	file.copy(rep, tmp)
+	rstudioapi::viewer(tmp)
+}

### 2019-05-19 util/map.R
Remove print from map function for render

@@ -20,6 +20,6 @@ ddr.map <- function(routes, autocenter = TRUE) {
 					  xlim = c(-20, 50), ylim = c(32, 70)) # Europe
 		}
 
-	print(map)
+	# print(map)
 	return(map)
 }

### 2019-05-19 report/report.Rmd
Add first figures to report

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+```
+
+```{r}
+bru <- 'EBBR'
+event <- c('2016-02-28', '2016-03-01')
+```
+```{r, fig.align='center'}
+ddr.map(data$real$routes[1:1000], autocenter = FALSE)
+```
+
+```{r, fig.align='center'}
+ddr.times(data$real$flights, bru, event)
 ```
@@
 
+- N flights vs. time with orig/dest at EBBR
+- N diverted flights vs. time with dest EBBR
+
 # Airports

### 2019-05-19 main.R
Remove tests and replace by render

@@ -3,7 +3,7 @@ day1 <- '2016/02/22'
 day2 <- '2016/03/02'
 
 # Define required packages
-requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj')
+requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj', 'dygraphs', 'rstudioapi')
 
 # Check requirements
 missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]
@@ -57,5 +57,6 @@ if (file.exists(file)) {
 	saveRDS(data, file)
 }
 
-# Test map plotting
-ddr.map(data$real$routes[1:1000], autocenter = FALSE)
+# Render report
+message('Rendering report...')
+ddr.render('report.Rmd')

### 2019-05-19 util/times.R
Change labels and add title function argument

@@ -1,4 +1,4 @@
-ddr.times <- function(flights, airport, event, onlydest = FALSE) {
+ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Traffic Evolution') {
 	cols <- c('date', 'N')
 	flightsWithDest <- flights[dest == airport, .N,  by = as.Date(landing)]
 	setnames(flightsWithDest, cols)
@@ -13,26 +13,26 @@ ddr.times <- function(flights, airport, event, onlydest = FALSE) {
 
 	if (!onlydest) {
 		dygraph(allFlights,
-				main = paste('Air Traffic Evolution at', airport),
+				main = paste(title, 'at', airport),
 				ylab = 'Number of Flights',
 				width = '100%') %>%
-			dySeries('N.x', label = paste(airport, 'as Origin')) %>%
-			dySeries('N.y', label = paste(airport, 'as Destination')) %>%
+			dySeries('N.x', label = paste(airport, 'Departures')) %>%
+			dySeries('N.y', label = paste(airport, 'Arrivals')) %>%
 			dyRangeSelector(dateWindow = event) %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%
-			dyLegend(width = '500') %>%
+			dyLegend(width = '450') %>%
 			dyEvent(event[1], 'Disruption begins', labelLoc = 'bottom') %>%
 			dyEvent(event[2], 'Disruption ends', labelLoc = 'bottom')
 	} else {
 		dygraph(allFlights,
-				main = paste('Air Traffic Evolution at', airport),
-				ylab = paste('Number of Flights')) %>%
-			dySeries('N', label = paste(airport, 'as Destination')) %>%
+				main = paste(title, 'at', airport),
+				ylab = paste('Number of Flights'),
+				width = '100%') %>%
+			dySeries('N', label = paste(airport, 'Arrivals')) %>%
 			dyRangeSelector(dateWindow = event) %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%
-			dyLegend(show = 'always', width = '500') %>%
+			dyLegend(show = 'always', width = '260') %>%
 			dyEvent(event[1], 'Disruption begins', labelLoc = 'bottom') %>%
 			dyEvent(event[2], 'Disruption ends', labelLoc = 'bottom')
 	}
-
 }

### 2019-05-19 util/map.R
Fix coordinates representation!

@@ -1,8 +1,8 @@
 ddr.map <- function(routes, autocenter = TRUE) {
 	map <- ggplot(map_data('world'), aes(x = long, y = lat, group = group)) +
 		geom_polygon(fill = 'white', color = 'black', size = 0.2) +
-		geom_line(data = routes,
-				  aes(x = lon, y = lat, group = id, color = rainbow(routes[, .N])),
+		geom_path(data = routes,
+				  aes(x = lon, y = lat, group = id, color = id),
 				  lwd = 0.25) +
 		theme(legend.position = 'none',
 			axis.line = element_blank(), axis.ticks = element_blank(),

### 2019-05-19 analysis.R
Source file with data analysis code

@@ -1,0 +1,9 @@
+# Define event
+bru <- 'EBBR'
+event <- c('2016-03-22', '2016-04-03')
+
+# Find diverted flights
+div <- merge(data$real$routes[, .SD[.N], by = id, .SDcols = c('pt')],
+			 data$real$flights[, .(id, dest, as.Date(landing))], by = c('id'))
+div <- div[dest == bru & pt != dest & V3 == event[1]] # v3 = date
+div <- div[, c('dest', 'V3') := NULL]

### 2019-05-19 main.R

@@ -1,6 +1,6 @@
 # Input date range
 day1 <- '2016/02/22'
-day2 <- '2016/03/02'
+day2 <- '2016/03/27'
 
 # Define required packages
 requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj', 'dygraphs', 'rstudioapi')
@@ -56,6 +56,9 @@ if (file.exists(file)) {
 	message(paste('Saving to full file:', basename(file)))
 	saveRDS(data, file)
 }
+
+# Call analysis
+source('analysis.R')
 
 # Render report
 message('Rendering report...')

### 2019-05-19 report/report.Rmd
Add plot of diverted flights

@@
 
-```{r}
-bru <- 'EBBR'
-event <- c('2016-02-28', '2016-03-01')
-```
+# Air Traffic
+
 ```{r, fig.align='center'}
@@
 
-# Air Traffic
+On the day of the event, March 22, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport (EBBR).
+
+```{r, fig.align='center'}
+ddr.map(data$real$routes[id %in% div[, id]], autocenter = FALSE)
+```
+
+## TODO
 
@@
 
+The nearby airports to EBBR include:
+- EBCI (Brussels South Charleroi)
+- EBLF (Liège)
+- EHEH (Eindhoven)
+- EBAW (Antwerp)
+- LFQQ (Lille)
+
 # Airlines

### 2019-05-19 util/map.R
Fix grouping colors!

@@ -2,7 +2,7 @@ ddr.map <- function(routes, autocenter = TRUE) {
 	map <- ggplot(map_data('world'), aes(x = long, y = lat, group = group)) +
 		geom_polygon(fill = 'white', color = 'black', size = 0.2) +
 		geom_path(data = routes,
-				  aes(x = lon, y = lat, group = id, color = id),
+				  aes(x = lon, y = lat, group = id, color = colors(distinct = TRUE)[id %% 502 + 1]),
 				  lwd = 0.25) +
 		theme(legend.position = 'none',
 			axis.line = element_blank(), axis.ticks = element_blank(),

### 2019-05-20 util/times.R
Update timeseries event labels and selector window

@@ -18,21 +18,21 @@ ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Tr
 				width = '100%') %>%
 			dySeries('N.x', label = paste(airport, 'Departures')) %>%
 			dySeries('N.y', label = paste(airport, 'Arrivals')) %>%
-			dyRangeSelector(dateWindow = event) %>%
+			dyRangeSelector() %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%
 			dyLegend(width = '450') %>%
-			dyEvent(event[1], 'Disruption begins', labelLoc = 'bottom') %>%
-			dyEvent(event[2], 'Disruption ends', labelLoc = 'bottom')
+			dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+			dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
 	} else {
 		dygraph(allFlights,
 				main = paste(title, 'at', airport),
 				ylab = paste('Number of Flights'),
 				width = '100%') %>%
 			dySeries('N', label = paste(airport, 'Arrivals')) %>%
-			dyRangeSelector(dateWindow = event) %>%
+			dyRangeSelector() %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%
 			dyLegend(show = 'always', width = '260') %>%
-			dyEvent(event[1], 'Disruption begins', labelLoc = 'bottom') %>%
-			dyEvent(event[2], 'Disruption ends', labelLoc = 'bottom')
+			dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+			dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
 	}
 }

### 2019-05-20 analysis.R
Complete fuel burn data analysis and create summary 'info' table

@@ -1,5 +1,6 @@
 # Define event
 bru <- 'EBBR'
+coord <- c(50.90139, 4.484444) # lat, lon
 event <- c('2016-03-22', '2016-04-03')
 
 # Find diverted flights
@@ -7,3 +8,24 @@ div <- merge(data$real$routes[, .SD[.N], by = id, .SDcols = c('pt')],
 			 data$real$flights[, .(id, dest, as.Date(landing))], by = c('id'))
 div <- div[dest == bru & pt != dest & V3 == event[1]] # v3 = date
 div <- div[, c('dest', 'V3') := NULL]
+
+# Fuel burn
+fuelburn <- 1.7862 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
+info <- merge(data$real$flights[id %in% div[, id], .(id, airline, orig, cumdist)],
+			  data$plan$flights[id %in% div[, id], .(id, cumdist)],
+			  by = c('id'))
+info[, diff := cumdist.x - cumdist.y]
+info[, c('cumdist.x', 'cumdist.y') := NULL]
+info <- merge(div, info, by = c('id'))
+info[, fuel := fuelburn * diff]
+
+# Airlines
+airlines <- readRDS(ddr.path('airlines'))
+# airlines <- airlines[month == unique(airlines$month)[which.min(abs(unique(airlines$month) - as.Date(event[1], '%Y-%m-%d')))]]
+airlines <- airlines[month == paste0(format(as.Date(event[1], '%Y-%m-%d'), '%Y-%m'), '-01')]
+info <- merge(info, airlines[, .(ICAO, name)], by.x = c('airline'), by.y = c('ICAO'))
+info <- info[, airline := NULL]
+
+# Final table
+info <- info[, .(id, orig, pt, name, diff, fuel)]
+setnames(info, c('id', 'orig', 'dest', 'airline', 'diff', 'fuel'))

### 2019-05-20 main.R
Switch to full date range and add new requirements for more interactive plots

@@ -1,9 +1,10 @@
 # Input date range
 day1 <- '2016/02/22'
-day2 <- '2016/03/27'
+day2 <- '2016/04/22'
 
 # Define required packages
-requirements <- c('data.table', 'fasttime', 'ggplot2', 'maps', 'mapproj', 'dygraphs', 'rstudioapi')
+requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools',
+				  'maps', 'mapproj', 'dygraphs',  'DT', 'leaflet')
 
 # Check requirements
 missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]

### 2019-05-20 util/jsmap.R
Function to create interactive maps!

@@ -1,0 +1,30 @@
+ddr.jsmap <- function(routes, info, coord) {
+	map <- leaflet(width = '100%') %>%
+		addTiles() %>%
+		addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
+		# addProviderTiles(providers$CartoDB.Positron) %>%
+		# addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
+
+		# addMarkers(lat = ~lat, lng = ~lon)
+		addAwesomeMarkers(lat = coord[1], lng = coord[2], icon = awesomeIcons(icon = 'plane',
+																			  iconColor = '#ffffff',
+																			  library = 'fa',
+																			  markerColor = 'red')) %>%
+		flyTo(lat = coord[1], lng = coord[2], zoom = 4)
+
+	# addPolylines(data = routes, lat = ~lat, lng = ~lon, group = ~id)
+	for (idx in unique(routes[, id])) {
+		infox <- info[id == idx]
+		map <- map %>%
+			addPolylines(data = routes[id == idx], lat = ~lat, lng = ~lon,
+						 color = paste0('#', paste(as.hexmode(c(col2rgb(colors(distinct = TRUE)[idx %% 502 + 1]))),
+						 						   collapse = '')),
+						 weight = 2, opacity = 1,
+						 highlightOptions = highlightOptions(weight = 5),
+						 label = HTML(paste0('<strong>', infox[1, orig], ' &ndash; ', infox[1, dest], '</strong>',
+						 					 '<br>', infox[1, airline])))
+	}
+
+	# print(map)
+	return(map)
+}

### 2019-05-20 report/report.Rmd
Add paged info table and interactive map! (tabbed)

@@
 ```{r, fig.align='center'}
-ddr.map(data$real$routes[1:1000], autocenter = FALSE)
-```
-
-```{r, fig.align='center'}
 ddr.times(data$real$flights, bru, event)
@@
 
+## Air Traffic Evolution {.tabset .tabset-fade}
+
+### Static Map {-}
+
 ```{r, fig.align='center'}
 ddr.map(data$real$routes[id %in% div[, id]], autocenter = FALSE)
+```
+
+### Interactive Map {-}
+
+```{r, fig.align='center'}
+ddr.jsmap(data$real$routes[id %in% div[, id]], info, coord)
+```
+
+On average, airlines actually saved `r format(abs(info[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
+
+```{r, fig.align='center'}
+datatable(info[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
+	formatRound(columns = c('diff', 'fuel'), digits = 2)
 ```
@@
 The nearby airports to EBBR include:
+
 - EBCI (Brussels South Charleroi)
@@
 
+## Traffic in nearby airports
+
+Find # of flights in nearby airports
+
+## Delays in nearby airports
+
 # Airlines

### 2019-05-20 main.R
Output missing status messages

@@ -59,8 +59,12 @@ if (file.exists(file)) {
 }
 
 # Call analysis
+message('Performing data analysis...')
 source('analysis.R')
 
 # Render report
 message('Rendering report...')
 ddr.render('report.Rmd')
+
+# End
+message('Done.')

### 2019-05-20 analysis.R
Fix fuelburn average

@@ -10,7 +10,7 @@ div <- div[dest == bru & pt != dest & V3 == event[1]] # v3 = date
 div <- div[, c('dest', 'V3') := NULL]
 
 # Fuel burn
-fuelburn <- 1.7862 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
+fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
 info <- merge(data$real$flights[id %in% div[, id], .(id, airline, orig, cumdist)],
 			  data$plan$flights[id %in% div[, id], .(id, cumdist)],
 			  by = c('id'))

### 2019-05-21 report/report.Rmd
Fix report sections organization
Add animation function and gif to report

@@
 
+## Air Traffic Evolution
+
 ```{r, fig.align='center'}
@@
 
+## Diverted Flights {.tabset .tabset-fade}
+
 On the day of the event, March 22, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport (EBBR).
 
-## Air Traffic Evolution {.tabset .tabset-fade}
+```{r, fig.align='center'}
+datatable(info[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
+	formatRound(columns = c('diff', 'fuel'), digits = 2)
+```
+
+On average, airlines actually saved `r format(abs(info[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
 
@@
 
-On average, airlines actually saved `r format(abs(info[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
+### Animation {-}
 
-```{r, fig.align='center'}
-datatable(info[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
-	formatRound(columns = c('diff', 'fuel'), digits = 2)
+```{r, message=FALSE, results=FALSE}
+gifname <- 'diverted'
+giffile <- file.path('..', 'data', paste0(gifname, '.gif'))
+if (!file.exists(giffile)) {
+	ddr.animate(data$real$routes[id %in% div[, id]], giffile)
+}
 ```
+![Diverted Flights Map Animation](`r giffile`) \
 
@@
 
+- Statistics (mean + sigma) for delays on each airport per day
+
 # Airlines

### 2019-05-21 main.R

@@ -4,7 +4,7 @@ day2 <- '2016/04/22'
 
 # Define required packages
 requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools',
-				  'maps', 'mapproj', 'dygraphs',  'DT', 'leaflet')
+				  'maps', 'mapproj', 'dygraphs',  'DT', 'leaflet', 'gganimate')
 
 # Check requirements
 missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]

### 2019-05-21 util/animate.R

@@ -1,0 +1,27 @@
+ddr.animate <- function(routes, filename, autocenter = TRUE) {
+	map <- ggplot(map_data('world'), aes(x = long, y = lat, group = group)) +
+		geom_polygon(fill = 'white', color = 'black', size = 0.2) +
+		geom_point(data = routes,
+				   aes(x = lon, y = lat, group = id, color = colors(distinct = TRUE)[id %% 502 + 1])) +
+		transition_components(dt) +
+		labs(title = 'Time: {frame_time}') +
+		shadow_wake(wake_length = 0.02) +
+		theme(legend.position = 'none',
+			  axis.line = element_blank(), axis.ticks = element_blank(),
+			  axis.text.x = element_blank(), axis.text.y = element_blank(),
+			  axis.title.x = element_blank(), axis.title.y = element_blank(),
+			  panel.border = element_rect(fill = NA, color = 'black')) +
+		if (autocenter) {
+			coord_map('ortho',
+					  orientation = c(median(routes[, lat], na.rm = TRUE), median(routes[, lon], na.rm = TRUE), 0),
+					  xlim = c(min(routes[, lon], na.rm = TRUE) - 2, max(routes[, lon], na.rm = TRUE) + 2),
+					  ylim = c(min(routes[, lat], na.rm = TRUE) - 2, max(routes[, lat], na.rm = TRUE) + 2))
+		} else {
+			coord_map('ortho',
+					  orientation = c(50.8, 4.4, 0), # Brussels
+					  xlim = c(-20, 50), ylim = c(32, 70)) # Europe
+		}
+
+	animate(map, width = 800, height = 600, detail = 5)
+	anim_save(filename)
+}

### 2019-05-21 main.R
Extend date range up to 26/04

@@ -1,6 +1,6 @@
 # Input date range
 day1 <- '2016/02/22'
-day2 <- '2016/04/22'
+day2 <- '2016/04/26'
 
 # Define required packages
 requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools',

### 2019-05-21 main.R
Add 'xts' package to remove startup message from report

@@ -3,7 +3,7 @@ day1 <- '2016/02/22'
 day2 <- '2016/04/26'
 
 # Define required packages
-requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools',
+requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools', 'xts',
 				  'maps', 'mapproj', 'dygraphs',  'DT', 'leaflet', 'gganimate')
 
 # Check requirements

### 2019-05-21 analysis.R
Convert 'event' to Date and find 'UFO'-type flights info

@@ -1,31 +1,39 @@
 # Define event
-bru <- 'EBBR'
-coord <- c(50.90139, 4.484444) # lat, lon
-event <- c('2016-03-22', '2016-04-03')
+bru <- 'EBBR' # airport ICAO code
+coord <- c(50.90139, 4.484444) # location (lat, lon)
+event <- c('2016/03/22', '2016/04/03') # interval (closes, reopens)
 
-# Find diverted flights
-div <- merge(data$real$routes[, .SD[.N], by = id, .SDcols = c('pt')],
-			 data$real$flights[, .(id, dest, as.Date(landing))], by = c('id'))
-div <- div[dest == bru & pt != dest & V3 == event[1]] # v3 = date
-div <- div[, c('dest', 'V3') := NULL]
-
-# Fuel burn
-fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
-info <- merge(data$real$flights[id %in% div[, id], .(id, airline, orig, cumdist)],
-			  data$plan$flights[id %in% div[, id], .(id, cumdist)],
-			  by = c('id'))
-info[, diff := cumdist.x - cumdist.y]
-info[, c('cumdist.x', 'cumdist.y') := NULL]
-info <- merge(div, info, by = c('id'))
-info[, fuel := fuelburn * diff]
+# Convert to dates
+event <- as.Date(event, '%Y/%m/%d')
 
 # Airlines
 airlines <- readRDS(ddr.path('airlines'))
+airlines <- airlines[month == paste0(format(event[1], '%Y-%m'), '-01')]
 # airlines <- airlines[month == unique(airlines$month)[which.min(abs(unique(airlines$month) - as.Date(event[1], '%Y-%m-%d')))]]
-airlines <- airlines[month == paste0(format(as.Date(event[1], '%Y-%m-%d'), '%Y-%m'), '-01')]
-info <- merge(info, airlines[, .(ICAO, name)], by.x = c('airline'), by.y = c('ICAO'))
-info <- info[, airline := NULL]
 
-# Final table
-info <- info[, .(id, orig, pt, name, diff, fuel)]
-setnames(info, c('id', 'orig', 'dest', 'airline', 'diff', 'fuel'))
+# Fuel burn
+fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
+
+# Diverted flights
[…]

### 2019-05-22 report/report.Rmd
Add UFO flights tables and update TODO lists

@@
 
-## Air Traffic Evolution
+## Air Traffic Evolution {.tabset .tabset-fade}
 
@@
 ddr.times(data$real$flights, bru, event)
+```
+
+- Find average before/after (get % difference)
+
+Bizarrely, despite the airport supposedly being closed from March 22 to April 3, the number of flights during the days following the attack is non-zero. Let's take a closer look at these flights:
+
+### EBBR Departures {-}
+
+```{r}
+datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
+	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString')
+```
+
+### EBBR Arrivals {-}
+
+```{r}
+datatable(ufo[dest == bru, .(orig, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
+	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString')
 ```
@@
 ```{r, fig.align='center'}
-datatable(info[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
+datatable(div[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
 	formatRound(columns = c('diff', 'fuel'), digits = 2)
@@
 
-On average, airlines actually saved `r format(abs(info[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
+On average, airlines actually saved `r format(abs(div[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
 
@@
 ```{r, fig.align='center'}
-ddr.jsmap(data$real$routes[id %in% div[, id]], info, coord)
+ddr.jsmap(data$real$routes[id %in% div[, id]], div, coord)
 ```
@@
 
-## TODO
-
-- N flights vs. time with orig/dest at EBBR
-- N diverted flights vs. time with dest EBBR
-
 # Airports
@@
 
-Find # of flights in nearby airports
+- Find # of flights in nearby airports vs. time (group time series)
 
@@
 
-- Statistics (mean + sigma) for delays on each airport per day
+- Delays on each airport vs. time (per hour)
+- Boxplot (mean + sigma) comparing average delays on each airport (while EBBR closed)
+- Find average delay before/during/after
 
 # Airlines
+
+- Find from total flights the airlines that were most affected (from those that travel to EBBR)
+- Average delays (while EBBR closed) by airline

### 2019-05-22 report/report.Rmd
Hide title and subtitle from sidebar toc

@@
 title: '<header style="text-align: center">
-			<h1 style="font-weight: bold; font-size: 44px">2016 Brussels Bombings</h1>
-			<h2>Air Traffic Analysis</h2>
+			<h1 style="font-weight: bold; font-size: 44px" class="toc-ignore">2016 Brussels Bombings</h1>
+			<h2 class="toc-ignore">Air Traffic Analysis</h2>
 		</header>'

### 2019-05-22 report/report.Rmd
Add grouped time series plots for nearby airports

@@
 
-- Find # of flights in nearby airports vs. time (group time series)
+```{r, fig.align='center', fig.height=2.5}
+ddr.times(data$real$flights, 'EBCI', event, group = 'nearby')
+# ddr.times(data$real$flights, 'EBLF', event, group = 'nearby')
+ddr.times(data$real$flights, 'EHEH', event, group = 'nearby')
+ddr.times(data$real$flights, 'EBAW', event, group = 'nearby')
+ddr.times(data$real$flights, 'LFQQ', event, group = 'nearby')
+```
+
+- Find average before/after (get % difference)
 

### 2019-05-22 util/times.R

@@ -1,4 +1,4 @@
-ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Traffic Evolution') {
+ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Traffic Evolution', group = NULL) {
 	cols <- c('date', 'N')
 	flightsWithDest <- flights[dest == airport, .N,  by = as.Date(landing)]
 	setnames(flightsWithDest, cols)
@@ -15,23 +15,25 @@ ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Tr
 		dygraph(allFlights,
 				main = paste(title, 'at', airport),
 				ylab = 'Number of Flights',
+				group = group,
 				width = '100%') %>%
 			dySeries('N.x', label = paste(airport, 'Departures')) %>%
 			dySeries('N.y', label = paste(airport, 'Arrivals')) %>%
 			dyRangeSelector() %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%
-			dyLegend(width = '450') %>%
+			dyLegend(width = '330') %>%
 			dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
 			dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
 	} else {
 		dygraph(allFlights,
 				main = paste(title, 'at', airport),
 				ylab = paste('Number of Flights'),
+				group = group,
 				width = '100%') %>%
 			dySeries('N', label = paste(airport, 'Arrivals')) %>%
 			dyRangeSelector() %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%
-			dyLegend(show = 'always', width = '260') %>%
+			dyLegend(show = 'always', width = '240') %>%
 			dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
 			dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
 	}

### 2019-05-22 util/times.R
Minor formatting fix

@@ -1,10 +1,10 @@
 ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Traffic Evolution', group = NULL) {
 	cols <- c('date', 'N')
-	flightsWithDest <- flights[dest == airport, .N,  by = as.Date(landing)]
+	flightsWithDest <- flights[dest == airport, .N, by = as.Date(landing)]
 	setnames(flightsWithDest, cols)
 
 	if (!onlydest) {
-		flightsWithOrig <- flights[orig == airport, .N,  by = as.Date(takeoff)]
+		flightsWithOrig <- flights[orig == airport, .N, by = as.Date(takeoff)]
 		setnames(flightsWithOrig, cols)
 		allFlights <- merge(flightsWithOrig, flightsWithDest, by = c('date'))
 	} else {

### 2019-05-22 report/report.Rmd
Analyze before/after event average flights, boxplot in report

@@
 
-- Find average before/after (get % difference)
+- Traffic before the attacks: `r floor(before[, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+- Traffic after the attacks: `r floor(after[, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+
+Even one month after the event, the traffic is still lower.
+Overall, the traffic is reduced in a `r format(abs((after[, mean(N.x + N.y)] - before[, mean(N.x + N.y)]) / before[, mean(N.x + N.y)]) * 100, digits = 3)`% following the tragic events.
+Further details can be extracted from the figure below:
+
+```{r, fig.align='center'}
+ggplot(data = NULL, aes(y = N.x + N.y)) +
+	geom_boxplot(data = before, aes(x = 'Before')) +
+	stat_summary(data = before, aes(x = 'Before'), fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
+	geom_boxplot(data = after, aes(x = 'After')) +
+	stat_summary(data = after, aes(x = 'After'), fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
+	scale_x_discrete(limits = c('Before', 'After')) +
+	labs(title = paste('Average Daily Flights at', bru), x = NULL, y = 'Number of Flights') +
+	theme(plot.title = element_text(hjust = 0.5))
+```
 

### 2019-05-22 analysis.R

@@ -37,3 +37,13 @@ ufo <- merge(ufo, airlines[, .(ICAO, name)], by.x = c('airline'), by.y = c('ICAO
 ufo[, airline := NULL]
 ufo <- ufo[, .(id, orig, dest, name, aircraft, takeoff, landing)]
 setnames(ufo, c('id', 'orig', 'dest', 'airline', 'aircraft', 'takeoff', 'landing'))
+
+# Before/After
+# mind <- as.Date(paste(format(event[1], '%Y'), as.integer(format(event[1], '%m')) - 1, format(event[1], '%d'),
+# 					  sep = '-'))
+before <- merge(data$real$flights[orig == bru & as.Date(takeoff) < event[1], .N, by = as.Date(takeoff)],
+				data$real$flights[dest == bru & as.Date(landing) < event[1], .N, by = as.Date(landing)],
+				by = c('as.Date'))
+after <- merge(data$real$flights[orig == bru & as.Date(takeoff) > event[2], .N, by = as.Date(takeoff)],
+			   data$real$flights[dest == bru & as.Date(landing) > event[2], .N, by = as.Date(landing)],
+			   by = c('as.Date'))

### 2019-05-24 report/report.Rmd
Add before/after/during traffic info and plot in nearby airports

@@
 - EBCI (Brussels South Charleroi)
-- EBLF (Liège)
+- EBLG (Liège)
+- EBAW (Antwerp)
 - EHEH (Eindhoven)
-- EBAW (Antwerp)
 - LFQQ (Lille)
@@
 ```{r, fig.align='center', fig.height=2.5}
-ddr.times(data$real$flights, 'EBCI', event, group = 'nearby')
-# ddr.times(data$real$flights, 'EBLF', event, group = 'nearby')
-ddr.times(data$real$flights, 'EHEH', event, group = 'nearby')
-ddr.times(data$real$flights, 'EBAW', event, group = 'nearby')
-ddr.times(data$real$flights, 'LFQQ', event, group = 'nearby')
+ddr.times(data$real$flights, nearby[1], event, group = 'nearby')
+ddr.times(data$real$flights, nearby[1], event, group = 'nearby')
+ddr.times(data$real$flights, nearby[2], event, group = 'nearby')
+ddr.times(data$real$flights, nearby[4], event, group = 'nearby')
+ddr.times(data$real$flights, nearby[5], event, group = 'nearby')
 ```
 
-- Find average before/after (get % difference)
+```{r, fig.align='center'}
+ggplot(data = NULL, aes(y = N.x + N.y)) +
+	geom_boxplot(data = befores, aes(x = orig, color = 'before')) +
+	geom_boxplot(data = durings, aes(x = orig, color = 'during')) +
+	geom_boxplot(data = afters, aes(x = orig, color = 'after')) +
+	scale_color_manual(limits = c('before', 'during', 'after'), values = c('green', 'red', 'blue'),
+					   name = NULL, labels = c('Before the attack', paste('While', bru, 'closed'),
+					   						   paste('After', bru, 'reopens'))) +
+	labs(title = 'Average Daily Flights in Nearby Airports', x = NULL, y = 'Number of Flights') +
+	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+```
+
+```{r, results='asis'}
+for (airport in nearby) {
+	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', (afters[orig == airport, mean(N.x + N.y)] - befores[orig == airport, mean(N.x + N.y)]) / befores[orig == airport, mean(N.x + N.y)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', (durings[orig == airport, mean(N.x + N.y)] - befores[orig == airport, mean(N.x + N.y)]) / befores[orig == airport, mean(N.x + N.y)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+}
+```
 

### 2019-05-24 analysis.R

@@ -2,6 +2,7 @@
 bru <- 'EBBR' # airport ICAO code
 coord <- c(50.90139, 4.484444) # location (lat, lon)
 event <- c('2016/03/22', '2016/04/03') # interval (closes, reopens)
+nearby <- c('EBCI', 'EBLG', 'EBAW', 'EHEH', 'LFQQ') # nearby airports
 
 # Convert to dates
 event <- as.Date(event, '%Y/%m/%d')
@@ -47,3 +48,18 @@ before <- merge(data$real$flights[orig == bru & as.Date(takeoff) < event[1], .N,
 after <- merge(data$real$flights[orig == bru & as.Date(takeoff) > event[2], .N, by = as.Date(takeoff)],
 			   data$real$flights[dest == bru & as.Date(landing) > event[2], .N, by = as.Date(landing)],
 			   by = c('as.Date'))
+befores <- merge(data$real$flights[orig %in% nearby & as.Date(takeoff) < event[1], .N,
+								   by = .(as.Date(takeoff), orig)],
+				 data$real$flights[dest %in% nearby & as.Date(landing) < event[1], .N,
+				 				   by = .(as.Date(landing), dest)],
+				 by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))
+durings <- merge(data$real$flights[orig %in% nearby & as.Date(takeoff) >= event[1] & as.Date(takeoff) <= event[2],
+								   .N, by = .(as.Date(takeoff), orig)],
+				 data$real$flights[dest %in% nearby & as.Date(landing) >= event[1] & as.Date(landing) <= event[2],
+				 				   .N, by = .(as.Date(landing), dest)],
+				 by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))
+afters <- merge(data$real$flights[orig %in% nearby & as.Date(takeoff) > event[2], .N,
+								   by = .(as.Date(takeoff), orig)],
+				 data$real$flights[dest %in% nearby & as.Date(landing) > event[2], .N,
+				 				  by = .(as.Date(landing), dest)],
+				 by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))

### 2019-05-24 report/report.Rmd
Loop to multiple figs from array

@@
 
-```{r, fig.align='center', fig.height=2.5}
-ddr.times(data$real$flights, nearby[1], event, group = 'nearby')
-ddr.times(data$real$flights, nearby[1], event, group = 'nearby')
-ddr.times(data$real$flights, nearby[2], event, group = 'nearby')
-ddr.times(data$real$flights, nearby[4], event, group = 'nearby')
-ddr.times(data$real$flights, nearby[5], event, group = 'nearby')
+```{r, fig.align='center'}
+htmltools::tagList(lapply(nearby, function(airport) {
+	ddr.times(data$real$flights, airport, event, group = 'nearby', height = '250px')
+}))
 ```

### 2019-05-24 util/times.R

@@ -1,4 +1,5 @@
-ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Traffic Evolution', group = NULL) {
+ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Traffic Evolution',
+					  group = NULL, height = NULL) {
 	cols <- c('date', 'N')
 	flightsWithDest <- flights[dest == airport, .N, by = as.Date(landing)]
 	setnames(flightsWithDest, cols)
@@ -16,7 +17,7 @@ ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Tr
 				main = paste(title, 'at', airport),
 				ylab = 'Number of Flights',
 				group = group,
-				width = '100%') %>%
+				width = '100%', height = height) %>%
 			dySeries('N.x', label = paste(airport, 'Departures')) %>%
 			dySeries('N.y', label = paste(airport, 'Arrivals')) %>%
 			dyRangeSelector() %>%
@@ -29,7 +30,7 @@ ddr.times <- function(flights, airport, event, onlydest = FALSE, title = 'Air Tr
 				main = paste(title, 'at', airport),
 				ylab = paste('Number of Flights'),
 				group = group,
-				width = '100%') %>%
+				width = '100%', height = height) %>%
 			dySeries('N', label = paste(airport, 'Arrivals')) %>%
 			dyRangeSelector() %>%
 			dyOptions(colors = c('royalblue', 'coral')) %>%

### 2019-05-24 report/report.Rmd
Remove html code and move css to 'styles.css' file

@@
 ---
-title: '<header style="text-align: center">
-			<h1 style="font-weight: bold; font-size: 44px" class="toc-ignore">2016 Brussels Bombings</h1>
-			<h2 class="toc-ignore">Air Traffic Analysis</h2>
-		</header>'
-author: '<div style="text-align: center">
-			<span style="display: inline-block; width: 30%; margin: 0 5px">[Ramiro Balado Ordax](https://github.com/Qjammer)</span>
-			<span style="display: inline-block; width: 30%; margin: 0 5px">[Alejandro Hernández Rubio](https://github.com/Shyryuu)</span>
-			<span style="display: inline-block; width: 30%; margin: 0 5px">[Pol Mesalles Ripoll](https://github.com/polmes)</span>
-		</div>'
-# date: "DD/MM/YYYY"
+title: '2016 Brussels Bombings'
+subtitle: 'Air Traffic Analysis'
+author:
+- '[Ramiro Balado Ordax](https://github.com/Qjammer)'
+- '[Alejandro Hernández Rubio](https://github.com/Shyryuu)'
+- '[Pol Mesalles Ripoll](https://github.com/polmes)'
 output:
@@
         toc_float: true
+        css: styles.css
 ---

### 2019-05-24 report/report.Rmd
Simplified before/during/after data structures

@@
 
-- Traffic before the attacks: `r floor(before[, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
-- Traffic after the attacks: `r floor(after[, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+- Traffic before the attacks: `r floor(before[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+- Traffic after the attacks: `r floor(after[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
 
 Even one month after the event, the traffic is still lower.
-Overall, the traffic is reduced in a `r format(abs((after[, mean(N.x + N.y)] - before[, mean(N.x + N.y)]) / before[, mean(N.x + N.y)]) * 100, digits = 3)`% following the tragic events.
+Overall, the traffic is reduced in a `r format(abs((after[orig == bru, mean(N.x + N.y)] - before[orig == bru, mean(N.x + N.y)]) / before[orig == bru, mean(N.x + N.y)]) * 100, digits = 3)`% following the tragic events.
 Further details can be extracted from the figure below:
@@
 ggplot(data = NULL, aes(y = N.x + N.y)) +
-	geom_boxplot(data = before, aes(x = 'Before')) +
-	stat_summary(data = before, aes(x = 'Before'), fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
-	geom_boxplot(data = after, aes(x = 'After')) +
-	stat_summary(data = after, aes(x = 'After'), fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
+	geom_boxplot(data = before[orig == bru], aes(x = 'Before')) +
+	stat_summary(data = before[orig == bru], aes(x = 'Before'),
+				 fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
+	geom_boxplot(data = after[orig == bru], aes(x = 'After')) +
+	stat_summary(data = after[orig == bru], aes(x = 'After'),
+				 fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
 	scale_x_discrete(limits = c('Before', 'After')) +
@@
 ggplot(data = NULL, aes(y = N.x + N.y)) +
-	geom_boxplot(data = befores, aes(x = orig, color = 'before')) +
-	geom_boxplot(data = durings, aes(x = orig, color = 'during')) +
-	geom_boxplot(data = afters, aes(x = orig, color = 'after')) +
+	geom_boxplot(data = before[orig != bru], aes(x = orig, color = 'before')) +
+	geom_boxplot(data = during, aes(x = orig, color = 'during')) +
+	geom_boxplot(data = after[orig != bru], aes(x = orig, color = 'after')) +
 	scale_color_manual(limits = c('before', 'during', 'after'), values = c('green', 'red', 'blue'),
@@
 for (airport in nearby) {
-	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', (afters[orig == airport, mean(N.x + N.y)] - befores[orig == airport, mean(N.x + N.y)]) / befores[orig == airport, mean(N.x + N.y)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', (durings[orig == airport, mean(N.x + N.y)] - befores[orig == airport, mean(N.x + N.y)]) / befores[orig == airport, mean(N.x + N.y)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', (after[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', (during[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
 }

### 2019-05-24 analysis.R

@@ -10,7 +10,6 @@ event <- as.Date(event, '%Y/%m/%d')
 # Airlines
 airlines <- readRDS(ddr.path('airlines'))
 airlines <- airlines[month == paste0(format(event[1], '%Y-%m'), '-01')]
-# airlines <- airlines[month == unique(airlines$month)[which.min(abs(unique(airlines$month) - as.Date(event[1], '%Y-%m-%d')))]]
 
 # Fuel burn
 fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
@@ -39,27 +38,19 @@ ufo[, airline := NULL]
 ufo <- ufo[, .(id, orig, dest, name, aircraft, takeoff, landing)]
 setnames(ufo, c('id', 'orig', 'dest', 'airline', 'aircraft', 'takeoff', 'landing'))
 
-# Before/After
-# mind <- as.Date(paste(format(event[1], '%Y'), as.integer(format(event[1], '%m')) - 1, format(event[1], '%d'),
-# 					  sep = '-'))
-before <- merge(data$real$flights[orig == bru & as.Date(takeoff) < event[1], .N, by = as.Date(takeoff)],
-				data$real$flights[dest == bru & as.Date(landing) < event[1], .N, by = as.Date(landing)],
-				by = c('as.Date'))
-after <- merge(data$real$flights[orig == bru & as.Date(takeoff) > event[2], .N, by = as.Date(takeoff)],
-			   data$real$flights[dest == bru & as.Date(landing) > event[2], .N, by = as.Date(landing)],
-			   by = c('as.Date'))
-befores <- merge(data$real$flights[orig %in% nearby & as.Date(takeoff) < event[1], .N,
-								   by = .(as.Date(takeoff), orig)],
-				 data$real$flights[dest %in% nearby & as.Date(landing) < event[1], .N,
-				 				   by = .(as.Date(landing), dest)],
-				 by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))
-durings <- merge(data$real$flights[orig %in% nearby & as.Date(takeoff) >= event[1] & as.Date(takeoff) <= event[2],
-								   .N, by = .(as.Date(takeoff), orig)],
-				 data$real$flights[dest %in% nearby & as.Date(landing) >= event[1] & as.Date(landing) <= event[2],
-				 				   .N, by = .(as.Date(landing), dest)],
[…]

### 2019-05-25 report/report.Rmd
Analyze delays in BRU + nearby airports, plot results

@@
 					   						   paste('After', bru, 'reopens'))) +
+	scale_x_discrete(limits = nearby) +
 	labs(title = 'Average Daily Flights in Nearby Airports', x = NULL, y = 'Number of Flights') +
@@
 
-## Delays in nearby airports
+## Delays
 
-- Delays on each airport vs. time (per hour)
-- Boxplot (mean + sigma) comparing average delays on each airport (while EBBR closed)
-- Find average delay before/during/after
+```{r, fig.align='center'}
+htmltools::tagList(lapply(c(bru, nearby), function(icao) {
+	dygraph(delays[airport == icao, .(date, depdel, arrdel)],
+			main = paste('Average Daily Delays at', icao),
+			ylab = paste('Delay (seconds)'),
+			group = 'delay',
+			width = '100%', height = '250px') %>%
+		dySeries('depdel', label = paste('Delay in Departures')) %>%
+		dySeries('arrdel', label = paste('Delay in Arrivals')) %>%
+		dyRangeSelector() %>%
+		dyOptions(colors = c('royalblue', 'coral')) %>%
+		dyLegend(width = '380') %>%
+		dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+		dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+}))
+```
+
+```{r, fig.align='center'}
+ggplot(data = NULL, aes(y = depdel)) +
+	geom_boxplot(data = delays[date < event[1]], aes(x = airport, color = 'before')) +
+	geom_boxplot(data = delays[date >= event[1] & date <= event[2]], aes(x = airport, color = 'during')) +
+	geom_boxplot(data = delays[date > event[2]], aes(x = airport, color = 'after')) +
+	scale_color_manual(limits = c('before', 'during', 'after'), values = c('green', 'red', 'blue'),
+					   name = NULL, labels = c('Before the attack', paste('While', bru, 'closed'),
+					   						   paste('After', bru, 'reopens'))) +
+	scale_x_discrete(limits = c(bru, nearby)) +
+	labs(title = 'Average Daily Departure Delays', x = NULL, y = 'Delay (seconds)') +
+	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+```
+
+```{r, fig.align='center'}
+ggplot(data = NULL, aes(y = arrdel)) +
+	geom_boxplot(data = delays[date < event[1]], aes(x = airport, color = 'before')) +
+	geom_boxplot(data = delays[date >= event[1] & date <= event[2]], aes(x = airport, color = 'during')) +
+	geom_boxplot(data = delays[date > event[2]], aes(x = airport, color = 'after')) +
+	scale_color_manual(limits = c('before', 'during', 'after'), values = c('green', 'red', 'blue'),
+					   name = NULL, labels = c('Before the attack', paste('While', bru, 'closed'),
+					   						   paste('After', bru, 'reopens'))) +
+	scale_x_discrete(limits = c(bru, nearby)) +
+	labs(title = 'Average Daily Arrival Delays', x = NULL, y = 'Delay (seconds)') +
+	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+```
+
+```{r, results='asis'}
+for (icao in nearby) {
+	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+}
+```
 

### 2019-05-25 analysis.R

@@ -54,3 +54,19 @@ after <- merge(data$real$flights[orig %in% c(bru, nearby) & as.Date(takeoff) > e
 			   data$real$flights[dest %in% c(bru, nearby) & as.Date(landing) > event[2], .N,
 								 by = .(as.Date(landing), dest)],
 			   by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))
+
+# Delays
+depdel <- merge(data$real$flights[orig %in% c(bru, nearby), .(id, orig, takeoff)],
+				data$plan$flights[orig %in% c(bru, nearby), .(id, takeoff)], by = c('id'))
+depdel[, delay := as.numeric(takeoff.x - takeoff.y)]
+depdel[, c('id', 'takeoff.y') := NULL]
+depdel <- depdel[, round(mean(delay)), by = .(as.Date(takeoff.x), orig)]
+arrdel <- merge(data$real$flights[dest %in% c(bru, nearby), .(id, dest, landing)],
+				data$plan$flights[dest %in% c(bru, nearby), .(id, landing)], by = c('id'))
+arrdel[, delay := as.numeric(landing.x - landing.y)]
+arrdel[, c('id', 'landing.y') := NULL]
+arrdel <- arrdel[, round(mean(delay)), by = .(as.Date(landing.x), dest)]
+lapply(list(depdel, arrdel), setnames, c('date', 'airport', 'delay'))
+delays <- merge(depdel, arrdel, by = c('date', 'airport'))
+rm(depdel, arrdel)
+setnames(delays, c('delay.x', 'delay.y'), c('depdel', 'arrdel'))

### 2019-05-25 util/jsmap.R
Plot plan routes in separate layer of leaflet interactive map

@@ -1,4 +1,4 @@
-ddr.jsmap <- function(routes, info, coord) {
+ddr.jsmap <- function(routes, info, coord, plan = NULL) {
 	map <- leaflet(width = '100%') %>%
 		addTiles() %>%
 		addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
@@ -22,7 +22,27 @@ ddr.jsmap <- function(routes, info, coord) {
 						 weight = 2, opacity = 1,
 						 highlightOptions = highlightOptions(weight = 5),
 						 label = HTML(paste0('<strong>', infox[1, orig], ' &ndash; ', infox[1, dest], '</strong>',
-						 					 '<br>', infox[1, airline])))
+						 					 '<br>', infox[1, airline])),
+						 group = 'Real')
+	}
+
+	if (!is.null(plan)) {
+		for (idx in unique(plan[, id])) {
+			infox <- info[id == idx]
+			map <- map %>%
+				addPolylines(data = plan[id == idx], lat = ~lat, lng = ~lon,
+							 color = paste0('#', paste(as.hexmode(c(col2rgb(colors(distinct = TRUE)[idx %% 502 + 1]))),
+							 						  collapse = '')),
+							 weight = 2, opacity = 1,
+							 highlightOptions = highlightOptions(weight = 5),
+							 label = HTML(paste0('<strong>', infox[1, orig], ' &ndash; ', infox[1, dest], '</strong>',
+							 					'<br>', infox[1, airline])),
+							 group = 'Plan')
+		}
+
+		map <- map %>%
+			addLayersControl(baseGroups = c('Real', 'Plan'),
+							 options = layersControlOptions(collapsed = FALSE))
 	}
 
 	# print(map)

### 2019-05-25 report/report.Rmd
Plot plan routes in separate layer of leaflet interactive map
Add missing delay info from EBBR

@@
 ```{r, fig.align='center'}
-ddr.jsmap(data$real$routes[id %in% div[, id]], div, coord)
+ddr.jsmap(data$real$routes[id %in% div[, id]], div, coord, data$plan$routes[id %in% div[, id]])
 ```
@@
 ```{r, results='asis'}
-for (icao in nearby) {
+for (icao in c(bru, nearby)) {
 	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))

### 2019-05-25 report/report.Rmd
Fix date time locale

@@
 datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
-	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString')
+	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString', params = list('ca-ES'))
 ```

### 2019-05-25 report/report.Rmd
FIXED: values now properly follow their definitions

@@
 for (airport in nearby) {
-	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', (after[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', (during[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', (during[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', (after[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
 }
@@
 for (icao in c(bru, nearby)) {
-	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 + 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
 }

### 2019-05-25 report/report.Rmd

@@
 for (airport in nearby) {
-	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', (during[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', (after[orig == airport, mean(N.x + N.y)] - before[orig == airport, mean(N.x + N.y)]) / before[orig == airport, mean(N.x + N.y)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', during[orig == airport, mean(N.x + N.y)] / before[orig == airport, mean(N.x + N.y)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', after[orig == airport, mean(N.x + N.y)] / before[orig == airport, mean(N.x + N.y)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
 }
@@
 for (icao in c(bru, nearby)) {
-	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & date > event[2], mean(depdel)] - delays[airport == icao & date < event[1], mean(depdel)]) / delays[airport == icao & date < event[1], mean(depdel)] * 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', (delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] - delays[airport == icao & date < event[1], mean(arrdel)]) / delays[airport == icao & date < event[1], mean(arrdel)] * 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 + 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] / delays[airport == icao & date < event[1], mean(depdel)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(depdel)] / delays[airport == icao & date < event[1], mean(depdel)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
 }

### 2019-05-25 report/report.Rmd
Analysis of most affected airlines

--- 2019-05-25 report/report.Rmd
@@
 - Average delays (while EBBR closed) by airline
+
+```{r}
+datatable(fucked, options = list(order = list(list(6, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
+	formatPercentage(columns = c('bd_rel', 'ba_rel'), digits = 1)
+```
+
+```{r, fig.align='center'}
+dygraph(fuckedplot,
+		main = 'Most Affected Airlines',
+		ylab = 'Number of Flights',
+		width = '100%') %>%
+	dyRangeSelector() %>%
+	dyLegend(width = '400') %>%
+	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```

### 2019-05-25 analysis.R

@@ -10,6 +10,7 @@ event <- as.Date(event, '%Y/%m/%d')
 # Airlines
 airlines <- readRDS(ddr.path('airlines'))
 airlines <- airlines[month == paste0(format(event[1], '%Y-%m'), '-01')]
+airlines <- airlines[, name := trimws(name)]
 
 # Fuel burn
 fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
@@ -70,3 +71,27 @@ lapply(list(depdel, arrdel), setnames, c('date', 'airport', 'delay'))
 delays <- merge(depdel, arrdel, by = c('date', 'airport'))
 rm(depdel, arrdel)
 setnames(delays, c('delay.x', 'delay.y'), c('depdel', 'arrdel'))
+
+# Affected Airlines
+fucked <- data$real$flights[, .N, by = .(as.Date(takeoff), airline)]
+fucked <- merge(merge(fucked[as.Date < event[1], round(mean(N)), by = airline],
+					  fucked[as.Date >= event[1] & as.Date <= event[2], round(mean(N)), by = airline],
+					  by = c('airline')),
+				fucked[as.Date > event[2], round(mean(N)), by = airline],
+				by = c('airline'))
+fucked[airline %in% unique(data$real$flights[orig == bru | dest == bru, airline])]
+fucked <- merge(fucked, airlines[, .(ICAO, name)], by.x = c('airline'), by.y = c('ICAO'))
+setnames(fucked, c('icao', 'before', 'during', 'after', 'airline'))
+fucked <- fucked[, .(airline, icao, before, during, after)]
+fucked[, bd_abs := during - before]
+fucked[, ba_abs := after - before]
+fucked[, bd_rel := (during - before) / before]
+fucked[, ba_rel := (after - before) / before]
+
+# Most Affected Airlines
+mostfucked <- fucked[order(bd_abs), icao][1:5]
+fuckedplot <- data$real$flights[airline %in% mostfucked & as.Date(takeoff) >= day1 & as.Date(takeoff) <= day2,
+								.N, by = .(as.Date(takeoff), airline)]
+fuckedplot <- reshape(fuckedplot, timevar = 'airline', idvar = 'as.Date', direction = 'wide')
+fuckedplot[is.na(fuckedplot)] = 0 # deal with <NA> values
+setnames(fuckedplot, c('date', mostfucked))

### 2019-05-25 main.R
Remove extra whitespace

@@ -4,7 +4,7 @@ day2 <- '2016/04/26'
 
 # Define required packages
 requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools', 'xts',
-				  'maps', 'mapproj', 'dygraphs',  'DT', 'leaflet', 'gganimate')
+				  'maps', 'mapproj', 'dygraphs', 'DT', 'leaflet', 'gganimate')
 
 # Check requirements
 missing <- requirements[!(requirements %in% installed.packages()[, 'Package'])]

### 2019-05-26 main.R
Ensure uniqueness in flights tables by id

@@ -48,10 +48,10 @@ if (file.exists(file)) {
 	names(data) <- names
 
 	# Merge tables + Ensure uniqueness
-	data <- list(real = list(flights = unique(rbindlist(lapply(data, function(x) x$real$flights))),
+	data <- list(real = list(flights = unique(rbindlist(lapply(data, function(x) x$real$flights)), by = c('id')),
 							 routes = unique(rbindlist(lapply(data, function(x) x$real$routes)))),
-				 plan = list(flights = unique(rbindlist(lapply(data, function(x) x$plan$flights))),
-				 			routes = unique(rbindlist(lapply(data, function(x) x$plan$routes)))))
+				 plan = list(flights = unique(rbindlist(lapply(data, function(x) x$plan$flights)), by = c('id')),
+				 			 routes = unique(rbindlist(lapply(data, function(x) x$plan$routes)))))
 
 	# Save for next time
 	message(paste('Saving to full file:', basename(file)))

### 2019-05-26 report/report.Rmd
Add European traffic over Belgium gif
New options in animate to control # of frames in animation

@@
 
+```{r, message=FALSE, results=FALSE}
+euname <- 'belgium'
+eufile <- file.path('..', 'data', paste0(euname, '.gif'))
+if (!file.exists(eufile)) {
+	ddr.animate(data$real$routes[id %in% data$real$flights[as.Date(takeoff) == event[1], id]], eufile, nf = 1000, autocenter = FALSE)
+}
+```
+![European Traffic Map Animation](`r eufile`) \
+
 Bizarrely, despite the airport supposedly being closed from March 22 to April 3, the number of flights during the days following the attack is non-zero. Let's take a closer look at these flights:
@@
 ```{r, message=FALSE, results=FALSE}
-gifname <- 'diverted'
-giffile <- file.path('..', 'data', paste0(gifname, '.gif'))
-if (!file.exists(giffile)) {
-	ddr.animate(data$real$routes[id %in% div[, id]], giffile)
+divname <- 'diverted'
+divfile <- file.path('..', 'data', paste0(divname, '.gif'))
+if (!file.exists(divfile)) {
+	ddr.animate(data$real$routes[id %in% div[, id]], divfile)
 }
 ```
-![Diverted Flights Map Animation](`r giffile`) \
+![Diverted Flights Map Animation](`r divfile`) \
 
@@
 
-- Find from total flights the airlines that were most affected (from those that travel to EBBR)
-- Average delays (while EBBR closed) by airline
-
 ```{r}
@@
 	dyRangeSelector() %>%
-	dyLegend(width = '400') %>%
+	dyLegend(width = '360') %>%
 	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%

### 2019-05-26 analysis.R

@@ -15,6 +15,9 @@ airlines <- airlines[, name := trimws(name)]
 # Fuel burn
 fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-haul flights
 
+# All European Traffic
+eu <- data$real$flights[as.Date(takeoff) >= day1 & as.Date(takeoff) <= day2, .N, by = as.Date(takeoff)]
+
 # Diverted flights
 div <- merge(data$real$flights[, .(id, as.Date(landing), orig, dest)],
 			 data$real$routes[, .SD[.N], by = id, .SDcols = c('pt')], by = c('id'))

### 2019-05-26 util/animate.R

@@ -1,11 +1,11 @@
-ddr.animate <- function(routes, filename, autocenter = TRUE) {
+ddr.animate <- function(routes, filename, nf = 100, autocenter = TRUE) {
 	map <- ggplot(map_data('world'), aes(x = long, y = lat, group = group)) +
 		geom_polygon(fill = 'white', color = 'black', size = 0.2) +
 		geom_point(data = routes,
 				   aes(x = lon, y = lat, group = id, color = colors(distinct = TRUE)[id %% 502 + 1])) +
 		transition_components(dt) +
 		labs(title = 'Time: {frame_time}') +
-		shadow_wake(wake_length = 0.02) +
+		shadow_wake(wake_length = 0.02 * 100/nf) +
 		theme(legend.position = 'none',
 			  axis.line = element_blank(), axis.ticks = element_blank(),
 			  axis.text.x = element_blank(), axis.text.y = element_blank(),
@@ -19,9 +19,9 @@ ddr.animate <- function(routes, filename, autocenter = TRUE) {
 		} else {
 			coord_map('ortho',
 					  orientation = c(50.8, 4.4, 0), # Brussels
-					  xlim = c(-20, 50), ylim = c(32, 70)) # Europe
+					  xlim = c(-2, 10), ylim = c(48, 54)) # Belgium
 		}
 
-	animate(map, width = 800, height = 600, detail = 5)
+	animate(map, width = 800, height = 600, detail = 5, nframes = nf)
 	anim_save(filename)
 }

### 2019-05-26 report/report.Rmd

--- 2019-05-25 report/report.Rmd
@@
 
-- Find from total flights the airlines that were most affected (from those that travel to EBBR)
-- Average delays (while EBBR closed) by airline
+```{r}
+datatable(fucked, options = list(order = list(list(6, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
+	formatPercentage(columns = c('bd_rel', 'ba_rel'), digits = 1)
+```
+
+```{r, fig.align='center'}
+dygraph(fuckedplot,
+		main = 'Most Affected Airlines',
+		ylab = 'Number of Flights',
+		width = '100%') %>%
+	dyRangeSelector() %>%
+	dyLegend(width = '360') %>%
+	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```

### 2019-05-26 main.R
Switch to final, fixed, and most complete dataset

@@ -1,6 +1,6 @@
 # Input date range
 day1 <- '2016/02/22'
-day2 <- '2016/04/26'
+day2 <- '2016/05/03'
 
 # Define required packages
 requirements <- c('data.table', 'fasttime', 'ggplot2', 'rstudioapi', 'htmltools', 'xts',

### 2019-05-26 report/report.Rmd
Add LEBL traffic evolution for comparison
Add European air traffic evolution plot and averages

@@
 
+Comparing the traffic evolution to another European airport, like Barcelona (IATA: BCN, ICAO: LEBL)...
+
+```{r, fig.align='center'}
+ddr.times(data$real$flights, 'LEBL', event) # for comparison: BCN
+```
+
 Even one month after the event, the traffic is still lower.
@@
 ```
+
+```{r, fig.align='center'}
+dygraph(eu, main = 'European Air Traffic Evolution', ylab = 'Total Number of Flights', width = '100%') %>%
+	dyRangeSelector() %>%
+	dyLegend(width = '160') %>%
+	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```
+
+- Average traffic before: `r format(round(mean(eu[date < event[1], N])))` flights per day
+- Average traffic while `r bru` was closed: `r format(round(mean(eu[date >= event[1] & date <= event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date >= event[1] & date <= event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
+- Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
 

### 2019-05-26 analysis.R

@@ -17,6 +17,7 @@ fuelburn <- 6.1264 # fuel burn [kg/NM] average of A320's and B737's for short-ha
 
 # All European Traffic
 eu <- data$real$flights[as.Date(takeoff) >= day1 & as.Date(takeoff) <= day2, .N, by = as.Date(takeoff)]
+setnames(eu, c('date', 'N'))
 
 # Diverted flights
 div <- merge(data$real$flights[, .(id, as.Date(landing), orig, dest)],

### 2019-05-26 report/report.Rmd
Extract info from BCN data

--- 2019-05-26 report/report.Rmd
@@
 
-- Traffic before the attacks: `r floor(before[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
-- Traffic after the attacks: `r floor(after[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+- Traffic before the attacks: `r round(before[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+- Traffic after the attacks: `r round(after[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
+
+Comparing the traffic evolution to another European airport, like Barcelona (IATA: BCN, ICAO: LEBL)...
+
+```{r, fig.align='center'}
+ddr.times(data$real$flights, bcn, event)
+```
+
+- Average traffic before the attacks: `r round(before[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals)
+- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals, `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
+- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals, `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
 
@@
 ```
+
+```{r, fig.align='center'}
+dygraph(eu, main = 'European Air Traffic Evolution', ylab = 'Total Number of Flights', width = '100%') %>%
+	dyRangeSelector() %>%
+	dyLegend(width = '160') %>%
+	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```
+
+- Average traffic before the attacks: `r format(round(mean(eu[date < event[1], N])))` flights per day
+- Average traffic while `r bru` was closed: `r format(round(mean(eu[date >= event[1] & date <= event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date >= event[1] & date <= event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
+- Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
 
@@
 ggplot(data = NULL, aes(y = N.x + N.y)) +
-	geom_boxplot(data = before[orig != bru], aes(x = orig, color = 'before')) +
-	geom_boxplot(data = during, aes(x = orig, color = 'during')) +
-	geom_boxplot(data = after[orig != bru], aes(x = orig, color = 'after')) +
+	geom_boxplot(data = before[orig %in% nearby], aes(x = orig, color = 'before')) +
+	geom_boxplot(data = during[orig %in% nearby], aes(x = orig, color = 'during')) +
+	geom_boxplot(data = after[orig %in% nearby], aes(x = orig, color = 'after')) +
 	scale_color_manual(limits = c('before', 'during', 'after'), values = c('green', 'red', 'blue'),

### 2019-05-26 analysis.R

@@ -3,6 +3,7 @@ bru <- 'EBBR' # airport ICAO code
 coord <- c(50.90139, 4.484444) # location (lat, lon)
 event <- c('2016/03/22', '2016/04/03') # interval (closes, reopens)
 nearby <- c('EBCI', 'EBLG', 'EBAW', 'EHEH', 'LFQQ') # nearby airports
+bcn <- 'LEBL' # airport for comparison
 
 # Convert to dates
 event <- as.Date(event, '%Y/%m/%d')
@@ -44,19 +45,19 @@ ufo <- ufo[, .(id, orig, dest, name, aircraft, takeoff, landing)]
 setnames(ufo, c('id', 'orig', 'dest', 'airline', 'aircraft', 'takeoff', 'landing'))
 
 # Before/During/After
-before <- merge(data$real$flights[orig %in% c(bru, nearby) & as.Date(takeoff) < event[1], .N,
+before <- merge(data$real$flights[orig %in% c(bru, nearby, bcn) & as.Date(takeoff) < event[1], .N,
 								  by = .(as.Date(takeoff), orig)],
-				data$real$flights[dest %in% c(bru, nearby) & as.Date(landing) < event[1], .N,
+				data$real$flights[dest %in% c(bru, nearby, bcn) & as.Date(landing) < event[1], .N,
 								  by = .(as.Date(landing), dest)],
 				by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))
-during <- merge(data$real$flights[orig %in% nearby & as.Date(takeoff) >= event[1] & as.Date(takeoff) <= event[2],
+during <- merge(data$real$flights[orig %in% c(bru, nearby, bcn) & as.Date(takeoff) >= event[1] & as.Date(takeoff) <= event[2],
 								  .N, by = .(as.Date(takeoff), orig)],
-				data$real$flights[dest %in% nearby & as.Date(landing) >= event[1] & as.Date(landing) <= event[2],
+				data$real$flights[dest %in% c(bru, nearby, bcn) & as.Date(landing) >= event[1] & as.Date(landing) <= event[2],
 								  .N, by = .(as.Date(landing), dest)],
 				by.x = c('as.Date', 'orig'), by.y = c('as.Date', 'dest'))
-after <- merge(data$real$flights[orig %in% c(bru, nearby) & as.Date(takeoff) > event[2], .N,
[…]

### 2019-05-26 report/report.Rmd
Reorganize report structure
Remove sort column from DTs

@@
 
-## Air Traffic Evolution {.tabset .tabset-fade}
+## Brussels Airport {.tabset .tabset-fade}
 
@@
 - Traffic after the attacks: `r round(after[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
-
-Comparing the traffic evolution to another European airport, like Barcelona (IATA: BCN, ICAO: LEBL)...
-
-```{r, fig.align='center'}
-ddr.times(data$real$flights, bcn, event)
-```
-
-- Average traffic before the attacks: `r round(before[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals)
-- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals, `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
-- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals, `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
 
@@
 
-```{r, fig.align='center'}
-dygraph(eu, main = 'European Air Traffic Evolution', ylab = 'Total Number of Flights', width = '100%') %>%
-	dyRangeSelector() %>%
-	dyLegend(width = '160') %>%
-	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
-```
-
-- Average traffic before the attacks: `r format(round(mean(eu[date < event[1], N])))` flights per day
-- Average traffic while `r bru` was closed: `r format(round(mean(eu[date >= event[1] & date <= event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date >= event[1] & date <= event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
-- Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
-
-```{r, message=FALSE, results=FALSE}
-euname <- 'belgium'
-eufile <- file.path('..', 'data', paste0(euname, '.gif'))
-if (!file.exists(eufile)) {
-	ddr.animate(data$real$routes[id %in% data$real$flights[as.Date(takeoff) == event[1], id]], eufile, nf = 1000, autocenter = FALSE)
-}
-```
-![European Traffic Map Animation](`r eufile`) \
-
 Bizarrely, despite the airport supposedly being closed from March 22 to April 3, the number of flights during the days following the attack is non-zero. Let's take a closer look at these flights:
@@
 ```{r}
-datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
+datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
 	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString', params = list('ca-ES'))
@@
 ```{r}
-datatable(ufo[dest == bru, .(orig, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
+datatable(ufo[dest == bru, .(orig, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
 	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString')
@@
 ```{r, fig.align='center'}
-datatable(div[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
+datatable(div[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
 	formatRound(columns = c('diff', 'fuel'), digits = 2)
@@
 ![Diverted Flights Map Animation](`r divfile`) \
+
+## European Sky
+
+```{r, fig.align='center'}
+dygraph(eu, main = 'European Air Traffic Evolution', ylab = 'Total Number of Flights', width = '100%') %>%
+	dyRangeSelector() %>%
+	dyLegend(width = '160') %>%
+	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
+	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```
+
+- Average traffic before the attacks: `r format(round(mean(eu[date < event[1], N])))` flights per day
+- Average traffic while `r bru` was closed: `r format(round(mean(eu[date >= event[1] & date <= event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date >= event[1] & date <= event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
+- Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
+
+```{r, message=FALSE, results=FALSE}
+euname <- 'belgium'
+eufile <- file.path('..', 'data', paste0(euname, '.gif'))
+if (!file.exists(eufile)) {
+	ddr.animate(data$real$routes[id %in% data$real$flights[as.Date(takeoff) == event[1], id]], eufile, nf = 1000, autocenter = FALSE)
+}
+```
+![European Traffic Map Animation](`r eufile`) \
+
+Comparing the traffic evolution to another European airport, like Barcelona (IATA: BCN, ICAO: LEBL)...
+
+```{r, fig.align='center'}
+ddr.times(data$real$flights, bcn, event)
+```
+
+- Average traffic before the attacks: `r round(before[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals)
+- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals, `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
+- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals, `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
 
@@
 ```{r}
-datatable(fucked, options = list(order = list(list(6, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
+datatable(fucked, options = list(order = list(list(6, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
 	formatPercentage(columns = c('bd_rel', 'ba_rel'), digits = 1)

### 2019-05-26 report/report.Rmd
Started adding report text

@@
 ```
+
+# Introduction {-}
+
+On Tuesday the 22nd of March, 2016, three suicidal terrorist attacks struck Brussels.
+Two of the explosions happened at 7:58am inside the main terminal of the Zaventem international airport (ICAO: EBBR, IATA: BRU), while the last blast struck the Maelbeek metro station in the city center an hour later.
+
+Following the incident, the airport closed until the 3rd of April, when it reopened again under a few restrictions.
+Many european countries reacted, including Germany, by stepping up security measures at points of high risk.
+
+The attack, that took the lives of 32 people and injured 340 more, raised the terrorist threat levels to "Severe", which symbolized a high possibility of attack.
+
+This study presents the results of an in-depth analysis of the air activity before, during and after the incident, addressing the effects of the shutdown of Brussels’ only international airport, as well as the fear that the attack induced after it.
 
@@
 On average, airlines actually saved `r format(abs(div[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
+However, costs related to those diverted flights such as compensations to passengers strongly outweigh those related to fuel consumption.
+
+The maps below display the routes followed by these diverted flights.
 
@@
 - Average traffic before the attacks: `r format(round(mean(eu[date < event[1], N])))` flights per day
-- Average traffic while `r bru` was closed: `r format(round(mean(eu[date >= event[1] & date <= event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date >= event[1] & date <= event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
-- Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`%)
+- Average traffic while `r bru` was closed: `r format(round(mean(eu[date >= event[1] & date <= event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date >= event[1] & date <= event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`% with respect to before the attack)
+- Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`% with respect to before the attack)
 
@@
 - Average traffic before the attacks: `r round(before[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals)
-- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals, `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
-- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals, `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`%)
+- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals; `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`% with respect to before the attack)
+- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals; `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`% with respect to before the attack)
 
@@
 ```
+
+# Final Notes {-}
+
+The R code developed for this project and used to render this report is freely available on [GitHub](https://github.com/polmes/r-ddr) under an MIT license.
+
+During the development of this study, one of the main issues has been the unreliability of the DDR2 platform, generating incomplete data at times, or simply not allowing all of us to generate downloads, no matter the time.
+At least RStudio will be [better](https://github.com/rstudio/rstudio/pull/4809) soon.

### 2019-05-27 analysis.R
Convert delay data to minutes

@@ -64,12 +64,12 @@ after <- merge(data$real$flights[orig %in% c(bru, nearby, bcn) & as.Date(takeoff
 # Delays
 depdel <- merge(data$real$flights[orig %in% c(bru, nearby), .(id, orig, takeoff)],
 				data$plan$flights[orig %in% c(bru, nearby), .(id, takeoff)], by = c('id'))
-depdel[, delay := as.numeric(takeoff.x - takeoff.y)]
+depdel[, delay := as.numeric(takeoff.x - takeoff.y) / 60] # convert to minutes
 depdel[, c('id', 'takeoff.y') := NULL]
 depdel <- depdel[, round(mean(delay)), by = .(as.Date(takeoff.x), orig)]
 arrdel <- merge(data$real$flights[dest %in% c(bru, nearby), .(id, dest, landing)],
 				data$plan$flights[dest %in% c(bru, nearby), .(id, landing)], by = c('id'))
-arrdel[, delay := as.numeric(landing.x - landing.y)]
+arrdel[, delay := as.numeric(landing.x - landing.y) / 60] # convert to minutes
 arrdel[, c('id', 'landing.y') := NULL]
 arrdel <- arrdel[, round(mean(delay)), by = .(as.Date(landing.x), dest)]
 lapply(list(depdel, arrdel), setnames, c('date', 'airport', 'delay'))

### 2019-05-27 report/report.Rmd
Finish 'Final Notes' text
Convert delay data to minutes!

@@
 Following the incident, the airport closed until the 3rd of April, when it reopened again under a few restrictions.
-Many european countries reacted, including Germany, by stepping up security measures at points of high risk.
+Many European countries reacted, including Germany, by stepping up security measures at points of high risk.
 
@@
 
+On a European scale, this event had a moderate impact on the air traffic.
+Due to the fact that the Brussels airport closed, many flights were rerouted and many more were cancelled.
+The close proximity in time with the french ATC strike compounded these effects and made the impact significantly worse.
+
+In this section, the impact of this terrorist bombing on the European and belgian air traffic is studied.
+In order to understand the data presented below, a bit of context is needed.
+Firstly, the European air traffic is very cyclic.
+Most of the flights occur between monday and friday, while the weekend sees a significant drop of flight numbers.
+
+Secondly, March 20th signals the beginning of the Holy Week, with March 27th being Easter.
+This generates a significant increase of air traffic throughout Europe, most likely due to the combination of national holidays and good weather.
+
 ## Brussels Airport {.tabset .tabset-fade}
+
+As it is expected, the airport faced a severe decrease in air traffic, both arrivals and departures, after the airport was reopened again.
+The traffic steadily increased during the first weeks after the reopening, but at a slow pace.
+On May 1st, the departures hall, which had sustained the most damage, partially reopened and marked the return of the airport to regular activity.
 
@@
 
+Computing the mean from the data above, we obtain:
+
 - Traffic before the attacks: `r round(before[orig == bru, mean(N.x + N.y)])` average daily flights (including departures and arrivals)
@@
 
-Even one month after the event, the traffic is still lower.
 Overall, the traffic is reduced in a `r format(abs((after[orig == bru, mean(N.x + N.y)] - before[orig == bru, mean(N.x + N.y)]) / before[orig == bru, mean(N.x + N.y)]) * 100, digits = 3)`% following the tragic events.
-Further details can be extracted from the figure below:
+Further details of this comparison are presented in the figure below, where the outer violin plot represents the flight density, while the inner box plot quantifies the distribution quartiles and the mean (diamond symbol):
 
@@
 ggplot(data = NULL, aes(y = N.x + N.y)) +
-	geom_boxplot(data = before[orig == bru], aes(x = 'Before')) +
+	geom_violin(data = before[orig == bru], aes(x = 'Before')) +
+	geom_boxplot(data = before[orig == bru], aes(x = 'Before'), width = 0.1) +
 	stat_summary(data = before[orig == bru], aes(x = 'Before'),
 				 fun.y = mean, geom = 'point', shape = 5, size = 2.5) +
-	geom_boxplot(data = after[orig == bru], aes(x = 'After')) +
+	geom_violin(data = after[orig == bru], aes(x = 'After')) +
+	geom_boxplot(data = after[orig == bru], aes(x = 'After'), width = 0.1) +
 	stat_summary(data = after[orig == bru], aes(x = 'After'),
@@
 
-Bizarrely, despite the airport supposedly being closed from March 22 to April 3, the number of flights during the days following the attack is non-zero. Let's take a closer look at these flights:
+Bizarrely, despite the airport supposedly being closed from March 22nd to April 3rd, the number of flights during the days following the terrorist attack is non-zero.
+Let's take a closer look at these flights:
 
@@
 
+Judging by the tables above, the flights during these days appear to be from regular airlines.
+Some of them might be from cargo flights, but for some reason some of them appear to be short-haul routes.
+
 ## Diverted Flights {.tabset .tabset-fade}
 
-On the day of the event, March 22, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport (EBBR).
+On the day of the bombing, March 22nd, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport (EBBR):
 
@@
 On average, airlines actually saved `r format(abs(div[, mean(fuel)]), digits = 3)` kg of fuel by diverting flights to airports closer to the origin point.
+These savings are so minimal that it had no appreciable effect either on the airlines costs or the environmental impact.
 However, costs related to those diverted flights such as compensations to passengers strongly outweigh those related to fuel consumption.
 
-The maps below display the routes followed by these diverted flights.
+The maps below detail the routes followed by these flights as they were diverted:
 
@@
 
+The total number of daily flights registered by Eurocontrol is plotted below:
+
 ```{r, fig.align='center'}
@@
 
+Overall, the effect over European air traffic is low, almost unnoticeable.
+This is due to the fact that the Brussels Airport does not control a large section of the European air traffic.
+This effect is even less noticeable taking into account the possibility of rerouting the planned flights through secondary airports.
+
+In fact, when analyzing the mean before, during and after the incident, we can see that the overall European volume of traffic actually increases:
+
 - Average traffic before the attacks: `r format(round(mean(eu[date < event[1], N])))` flights per day
@@
 - Average traffic after `r bru` reopens: `r format(round(mean(eu[date > event[2], N])))` flights per day (`r  sprintf('%+.1f', mean(eu[date > event[2], N]) / mean(eu[date < event[1], N]) * 100 - 100)`% with respect to before the attack)
+
+However, since the incident occurred around the Easter celebrations, the traffic was already expected to raise.
+This can be seen in airports such as Barcelona–El Prat Josep Tarradellas Airport (ICAO: LEBL, IATA: BCN), which showed almost no reaction to the incident and kept the expected evolution:
+
+```{r, fig.align='center'}
+ddr.times(data$real$flights, bcn, event)
+```
+
+In this case, the increase in traffic volume is significantly higher given the appeal of Barcelona:
+
+- Average traffic before the attacks: `r round(before[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals)
+- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals; `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`% with respect to before the attack)
+- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals; `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`% with respect to before the attack)
+
+Therefore, we can conclude that the overall impact of the terrorist attack is higher than what can be interpreted directly from the European air traffic evolution, since the number of flights would have increased more if it weren't for the generated insecurity.
+
+Finally, the animation below showcases the flights over Belgium and its surrounding countries over the course of the day of the tragic events.
+It can be seen that there is no significant change regarding the country's air space:
 
@@
 
-Comparing the traffic evolution to another European airport, like Barcelona (IATA: BCN, ICAO: LEBL)...
-
-```{r, fig.align='center'}
-ddr.times(data$real$flights, bcn, event)
-```
-
-- Average traffic before the attacks: `r round(before[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals)
-- Average traffic while `r bru` was closed: `r round(during[orig == bcn, mean(N.x + N.y)])` flights per day (including departures and arrivals; `r  sprintf('%+.1f', during[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`% with respect to before the attack)
-- Average traffic after `r bru` reopens: `r after[orig == bcn, mean(N.x + N.y)]` flights per day (including departures and arrivals; `r  sprintf('%+.1f', after[orig == bcn, mean(N.x + N.y)] / before[orig == bcn, mean(N.x + N.y)] * 100 - 100)`% with respect to before the attack)
-
 # Airports
 
-The nearby airports to EBBR include:
+In this section, the arrivals and departures before, during and after the studied event both in the Zaventem airport and others nearby are discussed.
+Points of notorious relevance like traffic density, tendency and delays will be analyzed.
 
-- EBCI (Brussels South Charleroi)
-- EBLG (Liège)
-- EBAW (Antwerp)
-- EHEH (Eindhoven)
-- LFQQ (Lille)
+The airports that have been considered to be close to Brussels and of relevance for this study are the following:
+
+- Brussels South Charleroi (ICAO: EBCI), 50km down from Brussels
+- Liège (ICAO: EBLG), eastern Belgium
+- Antwerp (ICAO: EBAW), northern Belgium
+- Eindhoven (ICAO: EHEH), southern Netherlands
+- Lille (ICAO: LFQQ), northern France
 
 ## Traffic in nearby airports
+
+To start, we will see how the traffic evolves in the five selected airports around Brussels, from one month before the attack to a month after the reopening of the airport:
 
@@
 ```
+
+The data from the plots above can be easily quantified by computing the mean:
+
+```{r, results='asis'}
+for (airport in nearby) {
+	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', during[orig == airport, mean(N.x + N.y)] / before[orig == airport, mean(N.x + N.y)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', after[orig == airport, mean(N.x + N.y)] / before[orig == airport, mean(N.x + N.y)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+}
+```
+
+Finally, in order to obtain statistically significant figures regarding the distribution, the box plots for each airport before, during and after the incident are drawn:
 
@@
 
-```{r, results='asis'}
-for (airport in nearby) {
-	cat(paste('-', airport, 'traffic:\n\t*', paste0(sprintf('%+.1f', during[orig == airport, mean(N.x + N.y)] / before[orig == airport, mean(N.x + N.y)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t*', paste0(sprintf('%+.1f', after[orig == airport, mean(N.x + N.y)] / before[orig == airport, mean(N.x + N.y)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
-}
-```
+We can see that right after the incident, there is a spike in both arrival and departure traffic density, due to the absorption of the traffic that Brussel’s main airport can’t take.
+The most notorious case is Liège (EBLG), where the traffic spiked by more than `r round(during[orig == 'EBLG', mean(N.x + N.y)] / before[orig == 'EBLG', mean(N.x + N.y)] * 100 - 100, digits = -2)`%.
+Following the opening of the main airport, the traffic slowly decreases as the Zaventem airport picks up the pace again.
+
+The one exception to this is Eindhoven (EHEH), which remains with its steady increase and appears unaffected by the incident.
 
 ## Delays
+
+In this section, the delays at Zaventem and nearby airports are studied. In the first few figures, the average daily delays as a function of time are plotted:
 
@@
 			main = paste('Average Daily Delays at', icao),
-			ylab = paste('Delay (seconds)'),
+			ylab = paste('Delay (minutes)'),
 			group = 'delay',
@@
 		dyOptions(colors = c('royalblue', 'coral')) %>%
-		dyLegend(width = '380') %>%
+		dyLegend(width = '350') %>%
 		dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
@@
 ```
+
+Note that there is a notable peak during the 20th of March, when an air traffic control strike over France paralyzed the french air space and caused numerous delays.
+Any flight that was scheduled to fly over France suffered from delays.
+In any case, that peak in the delays is not related to the incident being analyzed.
+
+As it can be seen in the graphs below and the means computed below, there is no relevant change in the delays of either arrivals or departures, during or after the incident.
+The data is mostly stochastic:
+
+```{r, results='asis'}
+for (icao in c(bru, nearby)) {
+	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] / delays[airport == icao & date < event[1], mean(depdel)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(depdel)] / delays[airport == icao & date < event[1], mean(depdel)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
+}
+```
+
+To sum up, the box plots below prove that neither arrival nor departure delays change in any noticeable way during or after the incident with respect to before:
 
@@
 	scale_x_discrete(limits = c(bru, nearby)) +
-	labs(title = 'Average Daily Departure Delays', x = NULL, y = 'Delay (seconds)') +
+	labs(title = 'Average Daily Departure Delays', x = NULL, y = 'Delay (minutes)') +
 	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
@@
 	scale_x_discrete(limits = c(bru, nearby)) +
-	labs(title = 'Average Daily Arrival Delays', x = NULL, y = 'Delay (seconds)') +
+	labs(title = 'Average Daily Arrival Delays', x = NULL, y = 'Delay (minutes)') +
 	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
@@
 
-```{r, results='asis'}
-for (icao in c(bru, nearby)) {
-	cat(paste('-', icao, 'delays:\n\t- Departures:\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & (date >= event[1] & date <= event[2]), mean(depdel)] / delays[airport == icao & date < event[1], mean(depdel)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(depdel)] / delays[airport == icao & date < event[1], mean(depdel)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n\t- Arrivals:\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & (date >= event[1] & date <= event[2]), mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 - 100), '%'), 'while', bru, 'was closed with respect to before the attack\n\t\t*', paste0(sprintf('%+.1f', delays[airport == icao & date > event[2], mean(arrdel)] / delays[airport == icao & date < event[1], mean(arrdel)] * 100 - 100), '%'), 'after', bru, 'reopens with respect to before the attack\n'))
-}
-```
+# Airlines
 
-# Airlines
+From an economical point of view, let’s look at the main airlines operating at Brussels Airport and analyze the evolution of flights before, during and after the incident.
+The table below summarizes these figures by presenting the average number of flights at each moment, as well as the absolute and relative differences:
 
@@
 ```
+
+Overall, there is a moderate loss in flight density during the incident, mostly due to a generalized feeling of insecurity.
+Brussels Airlines, as one could expect, was the airline that was struck harder by the events, dropping to half its average traffic.
+After the incidents were over and the Brussels Airport reopened, airlines gained traffic and even surpassed that of before: this is probably due to other events happening, such as Easter, that motivated traveling.
+
+The graphic below depicts the air traffic evolution of the top five airlines most affected by the terrorist bombing:
+
+- Brussels Airlines (ICAO: BEL), the flag carrier of Belgium (Star Alliance)
+- Scandinavian Airlines (ICAO: SAS), the flag carrier of Sweden, Norway and Denmark (Star Alliance)
+- Deutsche Lufthansa (ICAO: DLH), the largest German airline (Star Alliance)
+- Widerøes Flyveselskap (ICAO: WIF), the largest regional airline operating in the Nordic countries
+- Finnair (ICAO: FIN), the flag carrier of Finland (Oneworld)
 
@@
 
+It is interesting to note that the sudden drop in flights on the 27th of April for Brussels Airlines is due to a strike which affected all Lufthansa Group pilots, which Brussels Airlines belongs to.
+
 # Final Notes {-}
@@
 During the development of this study, one of the main issues has been the unreliability of the DDR2 platform, generating incomplete data at times, or simply not allowing all of us to generate downloads, no matter the time.
-At least RStudio will be [better](https://github.com/rstudio/rstudio/pull/4809) soon.
+
+At least RStudio will soon be [better](https://github.com/rstudio/rstudio/pull/4809) with our help.

### 2019-05-27 report/report.Rmd
Fix: air space -> airspace

@@
 Finally, the animation below showcases the flights over Belgium and its surrounding countries over the course of the day of the tragic events.
-It can be seen that there is no significant change regarding the country's air space:
+It can be seen that there is no significant change regarding the country's airspace:
 
@@
 
-Note that there is a notable peak during the 20th of March, when an air traffic control strike over France paralyzed the french air space and caused numerous delays.
+Note that there is a notable peak during the 20th of March, when an air traffic control strike over France paralyzed the french airspace and caused numerous delays.
 Any flight that was scheduled to fly over France suffered from delays.

### 2019-05-27 report/report.Rmd
Fix DT sorting

@@
 ```{r}
-datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
+datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(order = list(list(3, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
 	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString', params = list('ca-ES'))
@@
 ```{r}
-datatable(ufo[dest == bru, .(orig, airline, aircraft, takeoff, landing)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
+datatable(ufo[dest == bru, .(orig, airline, aircraft, takeoff, landing)], options = list(order = list(list(4, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
 	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString')
@@
 ```{r, fig.align='center'}
-datatable(div[, .(orig, dest, airline, diff, fuel)], options = list(dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
+datatable(div[, .(orig, dest, airline, diff, fuel)], options = list(order = list(list(1, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
 	formatRound(columns = c('diff', 'fuel'), digits = 2)
@@
 ```{r}
-datatable(fucked, options = list(order = list(list(6, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
+datatable(fucked, options = list(order = list(list(5, 'asc'), list(2, 'desc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
 	formatPercentage(columns = c('bd_rel', 'ba_rel'), digits = 1)

### 2019-07-12 util/render.R
Add support for multiple formats (i.e., html/pdf) in util.render function

@@ -1,6 +1,10 @@
-ddr.render <- function(file, dir = 'report') {
+ddr.render <- function(file, dir = 'report', format = 'html') {
 	rep <- rmarkdown::render(file.path(dir, file), quiet = TRUE)
-	tmp <- tempfile(file, fileext = '.html')
-	file.copy(rep, tmp)
-	rstudioapi::viewer(tmp)
+	if (format == 'html') {
+		tmp <- tempfile(file, fileext = paste0('.', format))
+		file.copy(rep, tmp)
+		rstudioapi::viewer(tmp)
+	} else if (format == 'pdf') {
+		system2(Sys.getenv('R_PDFVIEWER'), paste0(file.path(dir, tools::file_path_sans_ext(file)), '.pdf'))
+	}
 }

### 2019-07-12 report/paper.Rmd
Start report for paper

--- 2019-05-27 report/report.Rmd
@@
 author:
-- '[Ramiro Balado Ordax](https://github.com/Qjammer)'
-- '[Alejandro Hernández Rubio](https://github.com/Shyryuu)'
-- '[Pol Mesalles Ripoll](https://github.com/polmes)'
+- Ramiro Balado Ordax
+- Alejandro Hernández Rubio
+- Pol Mesalles Ripoll
 output:
-    rmarkdown::html_document:
-        theme: cosmo
+    pdf_document:
         highlight: tango
         number_sections: true
-        toc: true
-        toc_float: true
-        css: styles.css
+        keep_tex: true
+always_allow_html: true
 ---
@@
 
+## {- .toc-ignore}
+
 Judging by the tables above, the flights during these days appear to be from regular airlines.
@@
 
-### Static Map {-}
-
 ```{r, fig.align='center'}
@@
 ```
-
-### Interactive Map {-}
-
-```{r, fig.align='center'}
-ddr.jsmap(data$real$routes[id %in% div[, id]], div, coord, data$plan$routes[id %in% div[, id]])
-```
-
-### Animation {-}
-
-```{r, message=FALSE, results=FALSE}
-divname <- 'diverted'
-divfile <- file.path('..', 'data', paste0(divname, '.gif'))
-if (!file.exists(divfile)) {
-	ddr.animate(data$real$routes[id %in% div[, id]], divfile)
-}
-```
-![Diverted Flights Map Animation](`r divfile`) \
 
@@
 It can be seen that there is no significant change regarding the country's airspace:
-
-```{r, message=FALSE, results=FALSE}
-euname <- 'belgium'
-eufile <- file.path('..', 'data', paste0(euname, '.gif'))
-if (!file.exists(eufile)) {
-	ddr.animate(data$real$routes[id %in% data$real$flights[as.Date(takeoff) == event[1], id]], eufile, nf = 1000, autocenter = FALSE)
-}
-```
-![European Traffic Map Animation](`r eufile`) \
 

### 2019-07-12 util/render.R
Remove hardcoded file extensions

@@ -5,6 +5,6 @@ ddr.render <- function(file, dir = 'report', format = 'html') {
 		file.copy(rep, tmp)
 		rstudioapi::viewer(tmp)
 	} else if (format == 'pdf') {
-		system2(Sys.getenv('R_PDFVIEWER'), paste0(file.path(dir, tools::file_path_sans_ext(file)), '.pdf'))
+		system2(Sys.getenv('R_PDFVIEWER'), paste0(file.path(dir, tools::file_path_sans_ext(file)), '.', format))
 	}
 }

### 2019-07-12 main.R
Render paper report in pdf format by default now

@@ -64,7 +64,7 @@ source('analysis.R')
 
 # Render report
 message('Rendering report...')
-ddr.render('report.Rmd')
+ddr.render('paper.Rmd', format = 'pdf')
 
 # End
 message('Done.')

### 2019-07-12 report/paper.Rmd
Remove datatables from pdf report

@@
 
-### EBBR Departures {-}
-
-```{r}
-datatable(ufo[orig == bru, .(dest, airline, aircraft, takeoff, landing)], options = list(order = list(list(3, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Destination', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
-	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString', params = list('ca-ES'))
-```
-
-### EBBR Arrivals {-}
-
-```{r}
-datatable(ufo[dest == bru, .(orig, airline, aircraft, takeoff, landing)], options = list(order = list(list(4, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Airline', 'Aircraft', 'Departure', 'Arrival')) %>%
-	formatDate(columns = c('takeoff', 'landing'), method = 'toLocaleString')
-```
-
 ## {- .toc-ignore}
@@
 On the day of the bombing, March 22nd, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport (EBBR):
-
-```{r, fig.align='center'}
-datatable(div[, .(orig, dest, airline, diff, fuel)], options = list(order = list(list(1, 'asc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Origin', 'Destination', 'Airline', 'Distance Delta (NM)', 'Fuel Delta (kg)')) %>%
-	formatRound(columns = c('diff', 'fuel'), digits = 2)
-```
 
@@
 
-```{r}
-datatable(fucked, options = list(order = list(list(5, 'asc'), list(2, 'desc')), dom = 'tp', columnDefs = list(list(className = 'dt-center', targets = '_all'))), rownames = FALSE, colnames = c('Airline', 'ICAO Code', 'Flights Before', 'Flights During', 'Flights After', 'Delta During', 'Delta After', '% During', '% After')) %>%
-	formatPercentage(columns = c('bd_rel', 'ba_rel'), digits = 1)
-```
-
 Overall, there is a moderate loss in flight density during the incident, mostly due to a generalized feeling of insecurity.

### 2019-07-13 report/paper.Rmd
Replace dynamic (js) dygraphs with ggplots

@@
         keep_tex: true
+header-includes:
+- \usepackage{subfig}
 always_allow_html: true
@@
 
-```{r, fig.align='center'}
-ddr.times(data$real$flights, bru, event)
+```{r bru_traffic, fig.align='center', fig.cap=paste('Air Traffic Evolution at', bru)}
+ddr.stimes(data$real$flights, bru, event)
 ```
@@
 
-```{r, fig.align='center'}
+```{r bru_average, fig.align='center', fig.cap=paste('Average Daily Flights at', bru)}
 ggplot(data = NULL, aes(y = N.x + N.y)) +
@@
 	scale_x_discrete(limits = c('Before', 'After')) +
-	labs(title = paste('Average Daily Flights at', bru), x = NULL, y = 'Number of Flights') +
-	theme(plot.title = element_text(hjust = 0.5))
+	labs(x = NULL, y = 'Number of Flights')
 ```
@@
 
-```{r, fig.align='center'}
+```{r div_map, fig.align='center', fig.cap=paste('Map of all the diverted flights headed to', bru)}
 ddr.map(data$real$routes[id %in% div[, id]], autocenter = FALSE)
@@
 
-```{r, fig.align='center'}
-dygraph(eu, main = 'European Air Traffic Evolution', ylab = 'Total Number of Flights', width = '100%') %>%
-	dyRangeSelector() %>%
-	dyLegend(width = '160') %>%
-	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```{r eu_traffic, fig.align='center', fig.cap='European Air Traffic Evolution'}
+ggplot(data = eu) +
+	geom_line(aes(x = date, y = N, color = '')) +
+	scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+	geom_vline(xintercept = event, linetype = 2) +
+	labs(x = NULL, y = 'Total Number of Flights', color = NULL) +
+	annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
+	annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
+	theme(legend.position = 'none')
 ```
@@
 
-```{r, fig.align='center'}
-ddr.times(data$real$flights, bcn, event)
+```{r bcn_traffic, fig.align='center', fig.cap=paste('Air Traffic Evolution at', bcn)}
+ddr.stimes(data$real$flights, bcn, event)
 ```
@@
 
-```{r, fig.align='center'}
-htmltools::tagList(lapply(nearby, function(airport) {
-	ddr.times(data$real$flights, airport, event, group = 'nearby', height = '250px')
-}))
+```{r nearby_traffic, fig.align='center', fig.cap='Air Traffic Evolution in Nearby Airports', fig.subcap=nearby, fig.ncol=1}
+for (airport in nearby) {
+	print(ddr.stimes(data$real$flights, airport, event))
+}
 ```
@@
 
-```{r, fig.align='center'}
+```{r av_flights, fig.align='center', fig.cap='Average Daily Flights in Nearby Airports'}
 ggplot(data = NULL, aes(y = N.x + N.y)) +
@@
 	scale_x_discrete(limits = nearby) +
-	labs(title = 'Average Daily Flights in Nearby Airports', x = NULL, y = 'Number of Flights') +
-	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+	labs(x = NULL, y = 'Number of Flights') +
+	theme(legend.position = 'top')
 ```
@@
 
-```{r, fig.align='center'}
-htmltools::tagList(lapply(c(bru, nearby), function(icao) {
-	dygraph(delays[airport == icao, .(date, depdel, arrdel)],
-			main = paste('Average Daily Delays at', icao),
-			ylab = paste('Delay (minutes)'),
-			group = 'delay',
-			width = '100%', height = '250px') %>%
-		dySeries('depdel', label = paste('Delay in Departures')) %>%
-		dySeries('arrdel', label = paste('Delay in Arrivals')) %>%
-		dyRangeSelector() %>%
-		dyOptions(colors = c('royalblue', 'coral')) %>%
-		dyLegend(width = '350') %>%
-		dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-		dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
-}))
+```{r av_delays, fig.align='center', fig.cap='Average Daily Delays', fig.subcap=nearby, fig.ncol=1}
+for (icao in c(bru, nearby)) {
+	print(ggplot(data = delays[airport == icao, .(date, depdel, arrdel)], aes(x = date)) +
+		geom_line(aes(y = depdel, color = 'Delay in Departures')) +
+		geom_line(aes(y = arrdel, color = 'Delay in Arrivals')) +
+		scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+		geom_vline(xintercept = event, linetype = 2) +
+		labs(x = NULL, y = 'Delay (minutes)', color = NULL) +
+		annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
+		annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
+		theme(legend.position = 'top'))
+}
 ```
@@
 
-```{r, fig.align='center'}
+```{r av_dep, fig.align='center', fig.cap='Average Daily Departure Delays'}
 ggplot(data = NULL, aes(y = depdel)) +
@@
 	scale_x_discrete(limits = c(bru, nearby)) +
-	labs(title = 'Average Daily Departure Delays', x = NULL, y = 'Delay (minutes)') +
-	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+	labs(x = NULL, y = 'Delay (minutes)') +
+	theme(legend.position = 'top')
 ```
 
-```{r, fig.align='center'}
+```{r av_arr, fig.align='center', fig.cap='Average Daily Arrival Delays'}
 ggplot(data = NULL, aes(y = arrdel)) +
@@
 	scale_x_discrete(limits = c(bru, nearby)) +
-	labs(title = 'Average Daily Arrival Delays', x = NULL, y = 'Delay (minutes)') +
-	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+	labs(x = NULL, y = 'Delay (minutes)') +
+	theme(legend.position = 'top')
 ```
@@
 
-```{r, fig.align='center'}
-dygraph(fuckedplot,
-		main = 'Most Affected Airlines',
-		ylab = 'Number of Flights',
-		width = '100%') %>%
-	dyRangeSelector() %>%
-	dyLegend(width = '360') %>%
-	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```{r fucked, fig.align='center', fig.cap='Most Affected Airlines'}
+	ggplot(data = melt(fuckedplot[, 2:ncol(fuckedplot) := lapply(.SD, as.integer),
+								  .SDcols = 2:ncol(fuckedplot)], c('date'))) +
+		geom_line(aes(x = date, y = value, color = variable)) +
+		scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+		geom_vline(xintercept = event, linetype = 2) +
+		labs(x = NULL, y = 'Number of Flights', color = NULL) +
+		annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
+		annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
+		theme(legend.position = 'top')
 ```

### 2019-07-13 util/stimes.R

@@ -1,0 +1,34 @@
+ddr.stimes <- function(flights, airport, event, onlydest = FALSE) {
+	cols <- c('date', 'N')
+	flightsWithDest <- flights[dest == airport, .N, by = as.Date(landing)]
+	setnames(flightsWithDest, cols)
+
+	if (!onlydest) {
+		flightsWithOrig <- flights[orig == airport, .N, by = as.Date(takeoff)]
+		setnames(flightsWithOrig, cols)
+		allFlights <- merge(flightsWithOrig, flightsWithDest, by = c('date'))
+	} else {
+		allFlights <- flightsWithDest
+	}
+
+	if (!onlydest) {
+		ggplot(data = allFlights, aes(x = date)) +
+			geom_line(aes(y = N.x, color = paste(airport, 'Departures'))) +
+			geom_line(aes(y = N.y, color = paste(airport, 'Arrivals'))) +
+			scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+			geom_vline(xintercept = event, linetype = 2) +
+			labs(x = NULL, y = 'Number of Flights', color = NULL) +
+			annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
+			annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
+			theme(legend.position = 'top')
+	} else {
+		ggplot(data = allFlights, aes(x = date)) +
+			geom_line(aes(y = N, color = paste(airport, 'Arrivals'))) +
+			scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+			geom_vline(xintercept = event, linetype = 2) +
+			labs(x = NULL, y = 'Number of Flights', color = NULL) +
+			annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
+			annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
+			theme(legend.position = 'top')
+	}
+}

### 2019-07-13 report/paper.Rmd
Remove option to render html in pdf

@@
 - \usepackage{subfig}
-always_allow_html: true
 ---

### 2019-07-13 util/stimes.R
Added option to control interval date breaks

@@ -1,4 +1,4 @@
-ddr.stimes <- function(flights, airport, event, onlydest = FALSE) {
+ddr.stimes <- function(flights, airport, event, onlydest = FALSE, interval = '1 week') {
 	cols <- c('date', 'N')
 	flightsWithDest <- flights[dest == airport, .N, by = as.Date(landing)]
 	setnames(flightsWithDest, cols)
@@ -15,20 +15,20 @@ ddr.stimes <- function(flights, airport, event, onlydest = FALSE) {
 		ggplot(data = allFlights, aes(x = date)) +
 			geom_line(aes(y = N.x, color = paste(airport, 'Departures'))) +
 			geom_line(aes(y = N.y, color = paste(airport, 'Arrivals'))) +
-			scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+			scale_x_date(date_breaks = interval, date_labels = '%d/%m') +
 			geom_vline(xintercept = event, linetype = 2) +
 			labs(x = NULL, y = 'Number of Flights', color = NULL) +
-			annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
-			annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
+			annotate('text', x = event[1], label = 'Airport closes', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+			annotate('text', x = event[2], label = 'Airport reopens', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
 			theme(legend.position = 'top')
 	} else {
 		ggplot(data = allFlights, aes(x = date)) +
 			geom_line(aes(y = N, color = paste(airport, 'Arrivals'))) +
-			scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+			scale_x_date(date_breaks = interval, date_labels = '%d/%m') +
 			geom_vline(xintercept = event, linetype = 2) +
 			labs(x = NULL, y = 'Number of Flights', color = NULL) +
-			annotate('text', x = event[1], label = 'Airport closes', y = 10, angle = 90, vjust = -1, hjust = 0) +
-			annotate('text', x = event[2], label = 'Airport reopens', y = 10, angle = 90, vjust = -1, hjust = 0) +
[…]

### 2019-07-13 report/paper.Rmd

--- 2019-07-12 report/paper.Rmd
@@
 author:
-- Ramiro Balado Ordax
-- Alejandro Hernández Rubio
-- Pol Mesalles Ripoll
+- name: Ramiro Balado Ordax
+  affiliation: Universitat Politècnica de Catalunya
+- name: Alejandro Hernández Rubio
+  affiliation: Universitat Politècnica de Catalunya
+- name: Pol Mesalles Ripoll
+  affiliation: Universitat Politècnica de Catalunya
+abstract: |
+  In this study, the results from an in-depth data analysis of the effects the 2016 Brussels bombings had on the European sky are presented.
+Even though the attack is found to not have had a major impact on the European air traffic as a whole, it did result in major losses for Brussels Airport that lasted for weeks even after the international airport reopened.
+The consequences of this event for nearby airports are also studied, and a significant spike in operations is detected in most of them.
+Finally, the most affected airlines operating in Brussels are listed.
 output:
   pdf_document:
+    template: styles.tex
     highlight: tango
     number_sections: true
+    fig_caption: true
+    citation_package: biblatex
     keep_tex: true
-always_allow_html: true
+bibliography: references.bib
+graphics: true
+fontfamily: mathpazo
+fontsize: 11pt
+geometry: margin=2cm
+header-includes:
+- \usepackage{subfig}
+- \usepackage{float}
+- \usepackage{enumitem}
+- \setlist{nosep, before = \vspace{0pt}}
 ---
@@
 On Tuesday the 22nd of March, 2016, three suicidal terrorist attacks struck Brussels.
-Two of the explosions happened at 7:58am inside the main terminal of the Zaventem international airport (ICAO: EBBR, IATA: BRU), while the last blast struck the Maelbeek metro station in the city center an hour later.
+Two of the explosions happened at 7:58am inside the main terminal of the Zaventem international airport (ICAO: EBBR, IATA: BRU), while the last blast struck the Maelbeek metro station in the city center an hour later \cite{theguardian, cnn}.
 
@@
 This study presents the results of an in-depth analysis of the air activity before, during and after the incident, addressing the effects of the shutdown of Brussels’ only international airport, as well as the fear that the attack induced after it.
+All the data has been extracted from EUROCONTROL's DDR2 platform, which includes information from all planned and flown European flight routes \cite{ddr2}.
 
@@
 Firstly, the European air traffic is very cyclic.
-Most of the flights occur between monday and friday, while the weekend sees a significant drop of flight numbers.
+Most of the flights occur between Monday and Friday, while the weekend sees a significant drop of flight numbers.
 
@@
 
-As it is expected, the airport faced a severe decrease in air traffic, both arrivals and departures, after the airport was reopened again.
+As it is expected and can be seen in \autoref{fig:bru_traffic}, the airport faced a severe decrease in air traffic, both in terms of arrivals and departures, after it reopened again.
 The traffic steadily increased during the first weeks after the reopening, but at a slow pace.
 On May 1st, the departures hall, which had sustained the most damage, partially reopened and marked the return of the airport to regular activity.
+Computing the mean, we obtain:
 
-```{r, fig.align='center'}
-ddr.times(data$real$flights, bru, event)
+```{r bru_traffic, fig.align='center', fig.cap=paste('Air Traffic Evolution at', bru), fig.width=6, fig.height=4, fig.pos='p'}
+ddr.stimes(data$real$flights, bru, event)
 ```
-
-Computing the mean from the data above, we obtain:
 
@@
 Overall, the traffic is reduced in a `r format(abs((after[orig == bru, mean(N.x + N.y)] - before[orig == bru, mean(N.x + N.y)]) / before[orig == bru, mean(N.x + N.y)]) * 100, digits = 3)`% following the tragic events.
-Further details of this comparison are presented in the figure below, where the outer violin plot represents the flight density, while the inner box plot quantifies the distribution quartiles and the mean (diamond symbol):
+Further details of this comparison are presented in \autoref{fig:bru_average}, where the outer violin plot represents the flight density, while the inner box plot quantifies the distribution quartiles and the mean (diamond symbol).
 
-```{r, fig.align='center'}
+```{r bru_average, fig.align='center', fig.cap=paste('Average Daily Flights at', bru), fig.width=6, fig.height=4, fig.pos='p'}
 ggplot(data = NULL, aes(y = N.x + N.y)) +
@@
 Bizarrely, despite the airport supposedly being closed from March 22nd to April 3rd, the number of flights during the days following the terrorist attack is non-zero.
-Let's take a closer look at these flights:
-
-## {- .toc-ignore}
-
-Judging by the tables above, the flights during these days appear to be from regular airlines.
-Some of them might be from cargo flights, but for some reason some of them appear to be short-haul routes.
+By taking a closer look at said flights, they appear to be from regular airlines.
+Some of them might be from cargo flights, but for some reason some of them seem short-haul routes, though they might have been special flights with no passengers.
 
@@
 
-On the day of the bombing, March 22nd, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport (EBBR):
+On the day of the bombing, March 22nd, 2016, there were a total of `r div[, .N]` diverted flights that had intended to land on Brussels Airport.
 
@@
 
-The maps below detail the routes followed by these flights as they were diverted:
+\autoref{fig:div_map} showcases the detailed routes followed by these flights as they were diverted.
+In some cases, the airlines decided to return to the point of origin, while in others they had to land in different airports, usually close to EBBR.
 
-```{r, fig.align='center'}
+```{r div_map, fig.align='center', fig.cap=paste('Map of all the diverted flights headed to', bru), fig.height=3.4, fig.pos='H'}
 ddr.map(data$real$routes[id %in% div[, id]], autocenter = FALSE)
@@
 
-The total number of daily flights registered by Eurocontrol is plotted below:
-
-```{r, fig.align='center'}
-dygraph(eu, main = 'European Air Traffic Evolution', ylab = 'Total Number of Flights', width = '100%') %>%
-	dyRangeSelector() %>%
-	dyLegend(width = '160') %>%
-	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
-```
-
+The total number of daily flights registered by Eurocontrol is plotted in \autoref{fig:eu_traffic}.
 Overall, the effect over European air traffic is low, almost unnoticeable.
@@
 This effect is even less noticeable taking into account the possibility of rerouting the planned flights through secondary airports.
+In fact, when analyzing the mean before, during and after the incident, we can see that the overall European volume of traffic actually increases:
 
-In fact, when analyzing the mean before, during and after the incident, we can see that the overall European volume of traffic actually increases:
+```{r eu_traffic, fig.align='center', fig.cap='European Air Traffic Evolution', fig.width=6, fig.height=4, fig.pos='p'}
+ggplot(data = eu) +
+	geom_line(aes(x = date, y = N, color = '')) +
+	scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+	geom_vline(xintercept = event, linetype = 2) +
+	labs(x = NULL, y = 'Total Number of Flights', color = NULL) +
+	annotate('text', x = event[1], label = 'Airport closes', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+	annotate('text', x = event[2], label = 'Airport reopens', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+	theme(legend.position = 'none')
+```
 
@@
 However, since the incident occurred around the Easter celebrations, the traffic was already expected to raise.
-This can be seen in airports such as Barcelona–El Prat Josep Tarradellas Airport (ICAO: LEBL, IATA: BCN), which showed almost no reaction to the incident and kept the expected evolution:
+This can be seen in airports such as Barcelona–El Prat Josep Tarradellas Airport (ICAO: LEBL, IATA: BCN), a typical tourism destination, which showed almost no reaction to the incident and kept the expected evolution (see \autoref{fig:bcn_traffic}).
+In the case of Barcelona, the increase in traffic volume is significantly higher given the appeal of the city:
 
-```{r, fig.align='center'}
-ddr.times(data$real$flights, bcn, event)
+```{r bcn_traffic, fig.align='center', fig.cap=paste('Air Traffic Evolution at', bcn), fig.width=6, fig.height=4, fig.pos='p'}
+ddr.stimes(data$real$flights, bcn, event)
 ```
-
-In this case, the increase in traffic volume is significantly higher given the appeal of Barcelona:
 
@@
 
-Finally, the animation below showcases the flights over Belgium and its surrounding countries over the course of the day of the tragic events.
-It can be seen that there is no significant change regarding the country's airspace:
-
 # Airports
@@
 Points of notorious relevance like traffic density, tendency and delays will be analyzed.
-
 The airports that have been considered to be close to Brussels and of relevance for this study are the following:
@@
 
+\pagebreak
+
 ## Traffic in nearby airports
 
-To start, we will see how the traffic evolves in the five selected airports around Brussels, from one month before the attack to a month after the reopening of the airport:
+To start, in \autoref{fig:nearby_traffic} we see how the traffic evolves in the five selected airports around Brussels, from one month before the attack to a month after the reopening of the airport.
+The plotted data can also be easily quantified by computing the mean:
 
-```{r, fig.align='center'}
-htmltools::tagList(lapply(nearby, function(airport) {
-	ddr.times(data$real$flights, airport, event, group = 'nearby', height = '250px')
-}))
+```{r nearby_traffic, fig.align='center', fig.cap='Air Traffic Evolution in Nearby Airports', fig.subcap=nearby, fig.ncol=2, fig.width=3.25, fig.height=2.75, fig.pos='p'}
+for (airport in nearby) {
+	print(ddr.stimes(data$real$flights, airport, event, interval = '2 weeks'))
+}
 ```
-
-The data from the plots above can be easily quantified by computing the mean:
 
@@
 
-Finally, in order to obtain statistically significant figures regarding the distribution, the box plots for each airport before, during and after the incident are drawn:
+Finally, in order to obtain statistically significant figures regarding the distribution, box plots for each airport before, during and after the incident are drawn in \autoref{fig:av_flights}.
 
-```{r, fig.align='center'}
+```{r av_flights, fig.align='center', fig.cap='Average Daily Flights in Nearby Airports', fig.width=5, fig.height=3.4, fig.pos='H'}
 ggplot(data = NULL, aes(y = N.x + N.y)) +
@@
 Following the opening of the main airport, the traffic slowly decreases as the Zaventem airport picks up the pace again.
-
 The one exception to this is Eindhoven (EHEH), which remains with its steady increase and appears unaffected by the incident.
@@
 
-In this section, the delays at Zaventem and nearby airports are studied. In the first few figures, the average daily delays as a function of time are plotted:
-
-```{r, fig.align='center'}
-htmltools::tagList(lapply(c(bru, nearby), function(icao) {
-	dygraph(delays[airport == icao, .(date, depdel, arrdel)],
-			main = paste('Average Daily Delays at', icao),
-			ylab = paste('Delay (minutes)'),
-			group = 'delay',
-			width = '100%', height = '250px') %>%
-		dySeries('depdel', label = paste('Delay in Departures')) %>%
-		dySeries('arrdel', label = paste('Delay in Arrivals')) %>%
-		dyRangeSelector() %>%
-		dyOptions(colors = c('royalblue', 'coral')) %>%
-		dyLegend(width = '350') %>%
-		dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-		dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
-}))
-```
-
-Note that there is a notable peak during the 20th of March, when an air traffic control strike over France paralyzed the french airspace and caused numerous delays.
-Any flight that was scheduled to fly over France suffered from delays.
-In any case, that peak in the delays is not related to the incident being analyzed.
-
-As it can be seen in the graphs below and the means computed below, there is no relevant change in the delays of either arrivals or departures, during or after the incident.
+In this section, the delays at Zaventem and nearby airports are studied.
+No relevant change in either arrivals or departures, during or after the incident is found\footnote{Note that there is a notable peak during the 20th of March.
+An air traffic control strike paralyzed the French airspace and caused numerous delays, but these are not related to the incident being analyzed.}.
 The data is mostly stochastic:
@@
 
-To sum up, the box plots below prove that neither arrival nor departure delays change in any noticeable way during or after the incident with respect to before:
+In \autoref{fig:av_delays}, the average daily delays as a function of time are plotted.
+Then, box plots in Figures \ref{fig:av_dep} and \ref{fig:av_arr} prove that neither arrival nor departure delays change in any noticeable way during or after the incident with respect to before.
 
-```{r, fig.align='center'}
+```{r av_delays, fig.align='center', fig.cap='Average Daily Delays', fig.subcap=c(bru, nearby), fig.ncol=2, fig.width=3.25, fig.height=2.75, fig.pos='p'}
+for (icao in c(bru, nearby)) {
+	print(ggplot(data = delays[airport == icao, .(date, depdel, arrdel)], aes(x = date)) +
+		geom_line(aes(y = depdel, color = 'Delay in Departures')) +
+		geom_line(aes(y = arrdel, color = 'Delay in Arrivals')) +
+		scale_x_date(date_breaks = '2 weeks', date_labels = '%d/%m') +
+		geom_vline(xintercept = event, linetype = 2) +
+		labs(x = NULL, y = 'Delay (minutes)', color = NULL) +
+		annotate('text', x = event[1], label = 'Airport closes', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+		annotate('text', x = event[2], label = 'Airport reopens', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+		theme(legend.position = 'top'))
+}
+```
+
+```{r av_dep, fig.align='center', fig.cap='Average Daily Departure Delays', fig.width=6, fig.height=4, fig.pos='p'}
 ggplot(data = NULL, aes(y = depdel)) +
@@
 	scale_x_discrete(limits = c(bru, nearby)) +
-	labs(title = 'Average Daily Departure Delays', x = NULL, y = 'Delay (minutes)') +
-	theme(plot.title = element_text(hjust = 0.5), legend.position = 'bottom')
+	labs(x = NULL, y = 'Delay (minutes)') +
+	theme(legend.position = 'top')
 ```
 
-```{r, fig.align='center'}
+```{r av_arr, fig.align='center', fig.cap='Average Daily Arrival Delays', fig.width=6, fig.height=4, fig.pos='p'}
 ggplot(data = NULL, aes(y = arrdel)) +
@@
 
-From an economical point of view, let’s look at the main airlines operating at Brussels Airport and analyze the evolution of flights before, during and after the incident.
-The table below summarizes these figures by presenting the average number of flights at each moment, as well as the absolute and relative differences:
+From an economical point of view, it is interesting to look at the main airlines operating at Brussels Airport and analyze the evolution of their operations before, during and after the incident.
 
@@
 Brussels Airlines, as one could expect, was the airline that was struck harder by the events, dropping to half its average traffic.
-After the incidents were over and the Brussels Airport reopened, airlines gained traffic and even surpassed that of before: this is probably due to other events happening, such as Easter, that motivated traveling.
+After the incidents were over and the Brussels Airport reopened, airlines gained traffic and even surpassed that of before; this is probably due to other events happening, such as Easter, that motivated traveling.
 
-The graphic below depicts the air traffic evolution of the top five airlines most affected by the terrorist bombing:
+\autoref{fig:fucked} depicts the air traffic evolution of the top five airlines most affected by the terrorist bombing:
 
@@
 
-```{r, fig.align='center'}
-dygraph(fuckedplot,
-		main = 'Most Affected Airlines',
-		ylab = 'Number of Flights',
-		width = '100%') %>%
-	dyRangeSelector() %>%
-	dyLegend(width = '360') %>%
-	dyEvent(event[1], 'Airport closes', labelLoc = 'bottom') %>%
-	dyEvent(event[2], 'Airport reopens', labelLoc = 'bottom')
+```{r fucked, fig.align='center', fig.cap='Most Affected Airlines', fig.width=6, fig.height=4, fig.pos='H'}
+	ggplot(data = melt(fuckedplot[, 2:ncol(fuckedplot) := lapply(.SD, as.integer),
+								  .SDcols = 2:ncol(fuckedplot)], c('date'))) +
+		geom_line(aes(x = date, y = value, color = variable)) +
+		scale_x_date(date_breaks = '1 week', date_labels = '%d/%m') +
+		geom_vline(xintercept = event, linetype = 2) +
+		labs(x = NULL, y = 'Number of Flights', color = NULL) +
+		annotate('text', x = event[1], label = 'Airport closes', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+		annotate('text', x = event[2], label = 'Airport reopens', y = -Inf, angle = 90, vjust = -1, hjust = -0.1) +
+		theme(legend.position = 'top')
 ```
@@
 
-# Final Notes {-}
+# Conclusions {-}
 
-The R code developed for this project and used to render this report is freely available on [GitHub](https://github.com/polmes/r-ddr) under an MIT license.
+To summarize, there are two key aspects that had an impact in the European sky: the first, the shutdown and unavailability of Brussels main international airport caused a deficit in infrastructure capacity, leading to a considerable amount of flights diverted; the second, the fear induced by the terrorist attacks affected confidence in the security of airports, especially in the affected one, weakening its recovery after it was reopened again.
 
-During the development of this study, one of the main issues has been the unreliability of the DDR2 platform, generating incomplete data at times, or simply not allowing all of us to generate downloads, no matter the time.
+The former aspect is present mainly during the days following the attack.
+Secondary and nearby airports took the traffic load, and saw peaks in volume of traffic, increasing more than 600% in some cases (i.e., Liège).
 
-At least RStudio will soon be [better](https://github.com/rstudio/rstudio/pull/4809) with our help.
+The latter aspect is present after the airport was reopened.
+The recovery tendency was cut down by the lack of confidence and the temporary infrastructure put in place in Zaventem, which extended the time needed for a full recovery to more than a month after the bombings.
+Having said that, it is worth noting that the terror attack did not have a major impact at the European scale, but it is as well likely that many people cancelled their flights during a time of holidays due to the widespread fear.
+
+Lastly, note that the effects presented an increased difficulty in their study as the event happened at a critical time: right after an ACT strike in France, and during Holy Week and Easter, which changed the tendency of flights and had to be taken into account.

### 2019-07-13 main.R
Switch back to html report as default render

@@ -64,7 +64,7 @@ source('analysis.R')
 
 # Render report
 message('Rendering report...')
-ddr.render('paper.Rmd', format = 'pdf')
+ddr.render('report.Rmd') # for html, or: ddr.render('paper.Rmd', format = 'pdf') # for pdf
 
 # End
 message('Done.')
